### 2025-03-13 article/article.tex
add latex doc

\documentclass{article}
\usepackage{cite}
\usepackage{hyperref}

\title{Using graphs for advanced coverage arguments in autonomous driving.}
\author{xxx}
\date{\today}

\begin{document}

\maketitle

\section{Abstract}

\section{Introduction}

\section{exsting coverage approaches}

\section{Defining a traffic scene graph}

\section{Analysing a traffic scene graph from a coverage perspective}
\cite{newman2010networks}

\section{Application}

\subsection{Argoverse 2.0}

\subsection{Carla}

\cite{goodfellow2016deep}

\section{Summary}

\bibliographystyle{plain}
\bibliography{lit}

\end{document}

### 2025-03-13 carla/create_data_script_synchronous.py

@@ -32,6 +32,19 @@ j = 0
 world_map = world.get_map()
 map_g = MapGraph()
 map_g = map_g.create_from_carla_map(world_map)
+
+# map_g.graph.nodes(data=True)
+
+#topo = world_map.get_topology()
+
+#dir(topo[0][0])
+
+#wps = topo[0][0].next_until_lane_end(0.5)
+#lane_length = sum([wps[i].transform.location.distance(wps[i + 1].transform.location) for i in range(len(wps) - 1)])
+
+#topo[0][0].transform.location.distance(topo[0][1].transform.location)
+#len(wps)
+
 map_g.store_graph_to_file(f"carla/data/scene_{j}_{str(datetime.now().date())}_map_graph.pickle")
 
 # Get the world and set synchronous mode

### 2025-03-13 graph_creator/ActorGraph.py

@@ -62,6 +62,7 @@ class ActorGraph:
     def create_argoverse_actor_graphs(self, G_map):
         timestep_graphs = []
 
+        
         for t in range(len(next(iter(self.track_lane_dict.values())))):
             G_t = nx.MultiDiGraph()
 

### 2025-03-14 graph_creator/ActorGraph.py
change back some method names

@@ -55,11 +55,11 @@ class ActorGraph:
         instance.num_timesteps = len(scenario.timestamps_ns)
         instance.follow_vehicle_steps = follow_vehicle_steps
         instance.track_lane_dict = instance._create_track_lane_dict_argoverse(scenario)
-        instance.actor_graphs = instance.create_argoverse_actor_graphs(G_Map)
+        instance.actor_graphs = instance.create_actor_graphs(G_Map)
 
         return instance
 
-    def create_argoverse_actor_graphs(self, G_map):
+    def createactor_graphs(self, G_map):
         timestep_graphs = []
 
         

### 2025-03-14 carla/create_data_script_synchronous.py
small tweaks for actor graph

@@ -33,18 +33,6 @@ world_map = world.get_map()
 map_g = MapGraph()
 map_g = map_g.create_from_carla_map(world_map)
 
-# map_g.graph.nodes(data=True)
-
-#topo = world_map.get_topology()
-
-#dir(topo[0][0])
-
-#wps = topo[0][0].next_until_lane_end(0.5)
-#lane_length = sum([wps[i].transform.location.distance(wps[i + 1].transform.location) for i in range(len(wps) - 1)])
-
-#topo[0][0].transform.location.distance(topo[0][1].transform.location)
-#len(wps)
-
 map_g.store_graph_to_file(f"carla/data/scene_{j}_{str(datetime.now().date())}_map_graph.pickle")
 
 # Get the world and set synchronous mode
@@ -56,8 +44,6 @@ world.apply_settings(settings)
 # Get the traffic manager
 tm = client.get_trafficmanager(8000)  # Port 8000
 tm.set_synchronous_mode(True)  # Make TM sync with simulation
-
-
 
 
 # Get blueprint library

### 2025-03-14 graph_creator/ActorGraph.py

@@ -1,6 +1,7 @@
 from shapely.geometry import Point
 import networkx as nx
 import matplotlib.pyplot as plt
+from tqdm import tqdm
 
 class ActorGraph:
     # add opposite direction
@@ -40,13 +41,16 @@ class ActorGraph:
         if not len(lane_ids) == self.num_timesteps:
             raise ValueError(f"There are too many lane IDs for track {track.track_id}")
             
-        return lane_ids
-
+        return lane_ids 
+    
+    
     def _create_track_lane_dict_argoverse(self, scenario):
         track_lane_dict = {}
         for track in scenario.tracks:
             track_lane_dict[track.track_id] = self.find_lane_ids_for_track(track)
+    
         return track_lane_dict
+    
     
     @classmethod
     def from_argoverse_scenario(cls, scenario, G_Map, follow_vehicle_steps=3):
@@ -59,11 +63,38 @@ class ActorGraph:
 
         return instance
 
-    def createactor_graphs(self, G_map):
+    def _create_track_lane_dict_carla(self, scenario):
+        track_lane_dict = {}
+        actors = scenario.actor_id.unique().tolist()
+        for actor in actors:
+            mask = scenario.actor_id == actor
+            track_lane_dict[actor] = scenario[mask].road_lane_id.tolist()
+
+        return track_lane_dict
+
+
+    @classmethod
+    def from_carla_scenario(cls, scenario, G_Map, follow_vehicle_steps=3):
+        """
+        Args:
+            scenario: A pd dataframe with the following columns: 'track_id', 'timestep', 'x', 'y'
+            G_Map: A GraphMap object
+            follow_vehicle_steps: The maximum number of timesteps a vehicle can follow another vehicle
+        """
+        instance = cls()
+        instance.G_map = G_Map
+        instance.num_timesteps = scenario.timestamp.nunique()
+        instance.follow_vehicle_steps = follow_vehicle_steps
+        instance.track_lane_dict = instance._create_track_lane_dict_carla(scenario)
+        instance.actor_graphs = instance.create_actor_graphs(G_Map)
+
+        return instance
+
+    def create_actor_graphs(self, G_map):
         timestep_graphs = []
 
-        
-        for t in range(len(next(iter(self.track_lane_dict.values())))):
+        # das ist aber irgendwie eine ziemlich interessante Art die Anzahl an timestamps herauszufinden...
[…]

### 2025-04-02 graph_creator/ActorGraph.py
small updates

@@ -189,13 +189,90 @@ class ActorGraph:
 
         instance.track_lane_dict, instance.track_s_value_dict, instance.track_xyz_pos_dict, instance.track_speed_lon_dict = instance._create_track_lane_dict_carla_w_details(scenario)
 
-        instance.actor_graphs = instance.create_actor_graphs_w_details(G_Map, max_distance_lead_veh_m = 100, max_distance_neighbor_forward_m = 20, max_distance_neighbor_backward_m = 20, max_distance_opposite_m = 100)
+        instance.actor_graphs = instance.create_actor_graphs_w_details_2(G_Map, max_distance_lead_veh_m = 100, max_distance_neighbor_forward_m = 20, max_distance_neighbor_backward_m = 20, max_distance_opposite_m = 100)
+        instance.actor_components = {}
+
+        for key, value in instance.actor_graphs.items():
+            components = list(nx.weakly_connected_components(value))
+            subgraphs = [g.subgraph(c).copy() for c in components]
+            instance.actor_components[key] = subgraphs
 
         return instance
 
 
+    # def create_actor_graphs_w_details(self, G_map, max_distance_lead_veh_m, max_distance_neighbor_forward_m, max_distance_neighbor_backward_m, max_distance_opposite_m):
+    #     timestep_graphs = []
+
+    #     # das ist aber irgendwie eine ziemlich interessante Art die Anzahl an timestamps herauszufinden...
+        
+    #     #for t in tqdm(range(len(self.timestamps))):
+    #     for t in tqdm([0, 100, 200, 300, 400]):
+    #         G_t = nx.MultiDiGraph()
+
+    #         # Add nodes with attributes
+    #         for track_id, lane_ids in self.track_lane_dict.items():
+    #             if lane_ids[t] is not None:
+    #                 G_t.add_node(track_id, lane_id=lane_ids[t], 
+    #                              s = self.track_s_value_dict[track_id][t], 
+    #                              xyz = self.track_xyz_pos_dict[track_id][t], 
+    #                              lon_speed = self.track_speed_lon_dict[track_id][t])
+    #                 # Do we need to add for information about the track here? 
+
+
+    #         for track_id_A, lane_ids_A in self.track_lane_dict.items():
+
+    #             if lane_ids_A[t] is None:
+    #                 continue
+
+    #             for track_id_B, lane_ids_B in self.track_lane_dict.items():
+    #                 if track_id_A == track_id_B or lane_ids_B[t] is None:
+    #                     continue
[…]

### 2025-04-02 graph_creator/ActorGraph.py
small updates and opposite direction actor

@@ -298,6 +298,12 @@ class ActorGraph:
             max_distance_opposite_m=max_distance_opposite_veh_m,
         )
 
+        instance.actor_components = {}
+        for key, value in instance.actor_graphs.items():
+            components = list(nx.weakly_connected_components(value))
+            subgraphs = [value.subgraph(c).copy() for c in components]
+            instance.actor_components[key] = subgraphs
+
         return instance
 
     def _create_track_data_carla(self, scenario):
@@ -357,17 +363,28 @@ class ActorGraph:
                     # Do we need to add for information about the track here?
 
             # Add edges based on the conditions
-            for track_id_A, lane_ids_A in self.track_lane_dict.items():
+            keys = list(self.track_lane_dict.keys())
+            for i in range(len(keys) - 1):
+                track_id_A = keys[i]
+                lane_ids_A = self.track_lane_dict[keys[i]]
+                #for track_id_A, lane_ids_A in self.track_lane_dict.items():
+            
                 if lane_ids_A[t] is None:
                     continue
-                for track_id_B, lane_ids_B in self.track_lane_dict.items():
+
+                #for track_id_B, lane_ids_B in self.track_lane_dict.items():
+                for j in range(i + 1, len(keys)):
+                    track_id_B = keys[j]
+                    lane_ids_B = self.track_lane_dict[keys[j]]
+
                     if lane_ids_B[t] is None:
                         continue
-                    # Skip if we've already processed this pair
-                    if str(track_id_A) == str(track_id_B):
-                        continue
-                    if (track_id_B, track_id_A) in G_t.edges():
-                        continue
+                    # Skip if we've already processed this pair.
+                    # Should not happen with the above nested loops, but keep it to be on the safe side.
+                    # if str(track_id_A) == str(track_id_B):
+                    #    continue
+                    #if (track_id_B, track_id_A) in G_t.edges():
+                    #    continue
 
                     # Check for "following_lead" and "leading_vehicle"
                     if nx.has_path(G_map.graph, lane_ids_A[t], lane_ids_B[t]):
@@ -452,14 +469,29 @@ class ActorGraph:
                                     path_length=path_length,
                                 )
 
-                        # still need to add opposite direction.
+
+                        # fourth case: on opposite, directly next lane:
[…]

### 2025-04-04 graph_creator/ActorGraph.py
get it running with pydantic

@@ -8,7 +8,7 @@ import numpy as np
 
 
 class TrackData(BaseModel):
-    track_lane_dict: Dict[str, List[Optional[int]]] = Field(
+    track_lane_dict: Dict[str, List[Optional[str]]] = Field(
         description="Dictionary mapping track IDs to lists of lane IDs (can be None)"
     )
     track_s_value_dict: Dict[str, List[float]] = Field(description="Dictionary mapping track IDs to lists of s-values")
@@ -319,7 +319,7 @@ class ActorGraph:
             actor_id = str(actor)  # Convert to string
             mask = scenario.actor_id == actor
             # Convert lane IDs to integers
-            track_lane_dict[actor_id] = [int(lane_id) for lane_id in scenario[mask].road_lane_id.tolist()]
+            track_lane_dict[actor_id] = [lane_id for lane_id in scenario[mask].road_lane_id.tolist()]
             track_s_value_dict[actor_id] = scenario[mask].distance_from_lane_start.tolist()
             # Convert xyz coordinates to Shapely Points
             xyz_coords = scenario[mask].actor_location_xyz.tolist()
@@ -345,8 +345,8 @@ class ActorGraph:
         number_graphs = 10  # TODO: replace by delta_time .. Make sure time def in carla and argo is the same!
         timestep_delta = int(len(self.track_lane_dict) / number_graphs)
 
-        timestep_graphs = []
-
+        timestep_graphs = {}
+        print(len(self.track_lane_dict))
         for t in tqdm(range(0, len(self.track_lane_dict), timestep_delta)):
             G_t = nx.MultiDiGraph()
 
@@ -480,15 +480,15 @@ class ActorGraph:
                                 G_t.add_edge(track_id_A, track_id_B, edge_type='opposite_vehicle', path_length = path_length)
  
 
-            timestep_graphs.append(G_t)
+            timestep_graphs[t] = G_t
 
         return timestep_graphs
 
-    def visualize_actor_graph(self, t_idx, comp_idx, use_map_pos = True, node_size = 1600, save_path=None, graph_or_componentes = 'graph'):
+    def visualize_actor_graph(self, t_idx, comp_idx, use_map_pos = True, node_size = 1600, save_path=None, graph_or_component = 'graph'):
 
-        if graph_or_componentes == 'graph':
+        if graph_or_component == 'graph':
             G = self.actor_graphs[t_idx]
-        elif graph_or_componentes == 'component':
[…]

### 2025-04-04 carla/create_data_script_synchronous.py
some clean up and visualization try outs

@@ -19,17 +19,12 @@ os.getcwd()
 from graph_creator.MapGraph import MapGraph
 
 client = carla.Client("localhost", 2000)
-client.set_timeout(10.0)
+client.set_timeout(50.0)
 world = client.get_world()
 
 
 def to_2d(location):
     return (location.x, location.y)
-
-
-# clean_carla(world)
-
-client.load_world(random.choice(client.get_available_maps()))
 
 
 def get_t_coordinate(actor, world_map):
@@ -48,194 +43,137 @@ def get_t_coordinate(actor, world_map):
     return t
 
 
-j = 0
-
-# create the lane map graph and store it to file:
-world_map = world.get_map()
-map_g = MapGraph()
-map_g = map_g.create_from_carla_map(world_map)
-
-map_g.store_graph_to_file(f"carla/data/scene_{j}_{str(datetime.now().date())}_map_graph.pickle")
-
-xodr = world_map.to_opendrive()
-with open(f"carla/data/scene_{j}_{str(datetime.now().date())}_map.xodr", "w") as f:
-    f.write(xodr)
-
-# Get the world and set synchronous mode
-settings = world.get_settings()
-settings.synchronous_mode = True  # Enable sync mode
-settings.fixed_delta_seconds = 0.05  # Set fixed time step
-world.apply_settings(settings)
-
-# Get the traffic manager
-tm = client.get_trafficmanager(8000)  # Port 8000
-tm.set_synchronous_mode(True)  # Make TM sync with simulation
-
-
-# Get blueprint library
-blueprint_library = world.get_blueprint_library()
-
-# Spawn vehicles
-spawn_points = world.get_map().get_spawn_points()
-random.shuffle(spawn_points)
-
-
-for i in range(25):
-    # vehicle_bp = blueprint_library.filter("vehicle.*")[0]
-    spawn_point = spawn_points[i]
-    vehicle = world.spawn_actor(random.choice(blueprint_library.filter("vehicle.*")), spawn_point)
-    # Enable autopilot
-    vehicle.set_autopilot(True, tm.get_port())  # TM handles driving
-    # vehicles.append(vehicle)
-
-
-tracks = []
-n_steps = 500
-
-
-for i in tqdm(range(n_steps)):
-    world.tick()
-    for act in world.get_actors().filter("vehicle.*"):
-        # act = world.get_actors().filter("vehicle.*")[0]
-        row = {}
-        row["actor_id"] = act.id
-        row["actor_type"] = act.type_id
[…]

### 2025-04-05 graph_creator/ActorGraph.py
halbgares

@@ -16,13 +16,16 @@ class TrackData(BaseModel):
     track_speed_lon_dict: Dict[str, List[float]] = Field(
         description="Dictionary mapping track IDs to lists of longitudinal speeds"
     )
+    track_actor_type_dict: Dict[str, List[str]] = Field(description="Dictionary mapping track IDs to actor types.")
 
     @field_validator(
         "track_lane_dict",
         "track_s_value_dict",
         "track_xyz_pos_dict",
         "track_speed_lon_dict",
+        "track_actor_type_dict",
     )
+
     @classmethod
     def validate_list_lengths(cls, v: Dict[str, List]) -> Dict[str, List]:
         if not v:  # Skip validation if dictionary is empty
@@ -50,13 +53,13 @@ class TrackData(BaseModel):
         s_value_dict = self.track_s_value_dict
         xyz_dict = self.track_xyz_pos_dict
         speed_dict = self.track_speed_lon_dict
-
+        actor_dict = self.track_actor_type_dict
         # Get all unique actors across all dictionaries
-        all_actors = set(lane_dict.keys()) | set(s_value_dict.keys()) | set(xyz_dict.keys()) | set(speed_dict.keys())
+        all_actors = set(lane_dict.keys()) | set(s_value_dict.keys()) | set(xyz_dict.keys()) | set(speed_dict.keys() | set(actor_dict.keys()))
 
         # Check that all dictionaries have the same actors
         for actor in all_actors:
-            if not all(actor in d for d in [lane_dict, s_value_dict, xyz_dict, speed_dict]):
+            if not all(actor in d for d in [lane_dict, s_value_dict, xyz_dict, speed_dict, actor_dict]):
                 raise ValueError(f"Actor {actor} is missing in one or more dictionaries")
 
         # Check that all lists have the same length for each actor
@@ -66,6 +69,8 @@ class TrackData(BaseModel):
                 len(s_value_dict[actor]),
                 len(xyz_dict[actor]),
                 len(speed_dict[actor]),
+                len(actor_dict[actor]),
+                # As actor_dict isnot time indexed, it does not need to be checked here.,
             ]
             if not all(l == lengths[0] for l in lengths):
                 raise ValueError(f"Inconsistent list lengths for actor {actor}. Expected {lengths[0]}, got {lengths}")
@@ -240,6 +245,7 @@ class ActorGraph:
         instance.track_s_value_dict = track_data.track_s_value_dict
[…]

### 2025-04-05 graph_creator/ActorGraph.py
3/4 gares

@@ -334,7 +334,6 @@ class ActorGraph:
             track_xyz_pos_dict[actor_id] = [Point(x, y, z) for x, y, z in xyz_coords]
             track_speed_lon_dict[actor_id] = scenario[mask].actor_speed_lon.tolist()
             track_actor_type_dict[actor_id] = scenario[mask].actor_type.tolist()
-        print(track_actor_type_dict)
         # Create and return the Pydantic model
         return TrackData(
             track_lane_dict=track_lane_dict,

### 2025-04-07 graph_creator/ActorGraph.py
Changed actor type from List[str] to str.  Argoverse to pydantic

@@ -16,14 +16,13 @@ class TrackData(BaseModel):
     track_speed_lon_dict: Dict[str, List[float]] = Field(
         description="Dictionary mapping track IDs to lists of longitudinal speeds"
     )
-    track_actor_type_dict: Dict[str, List[str]] = Field(description="Dictionary mapping track IDs to actor types.")
+    track_actor_type_dict: Dict[str, str] = Field(description="Dictionary mapping track IDs to actor types.")
 
     @field_validator(
         "track_lane_dict",
         "track_s_value_dict",
         "track_xyz_pos_dict",
         "track_speed_lon_dict",
-        "track_actor_type_dict",
     )
 
     @classmethod
@@ -55,7 +54,7 @@ class TrackData(BaseModel):
         speed_dict = self.track_speed_lon_dict
         actor_dict = self.track_actor_type_dict
         # Get all unique actors across all dictionaries
-        all_actors = set(lane_dict.keys()) | set(s_value_dict.keys()) | set(xyz_dict.keys()) | set(speed_dict.keys() | set(actor_dict.keys()))
+        all_actors = set(lane_dict.keys()) | set(s_value_dict.keys()) | set(xyz_dict.keys()) | set(speed_dict.keys()) | set(actor_dict.keys())
 
         # Check that all dictionaries have the same actors
         for actor in all_actors:
@@ -69,8 +68,6 @@ class TrackData(BaseModel):
                 len(s_value_dict[actor]),
                 len(xyz_dict[actor]),
                 len(speed_dict[actor]),
-                len(actor_dict[actor]),
-                # As actor_dict isnot time indexed, it does not need to be checked here.,
             ]
             if not all(l == lengths[0] for l in lengths):
                 raise ValueError(f"Inconsistent list lengths for actor {actor}. Expected {lengths[0]}, got {lengths}")
@@ -113,7 +110,7 @@ class ActorGraph:
             if ii in timestep_list:
                 position = track.object_states[timestep_list.index(ii)].position
                 lane_id = self.find_lane_id_from_pos(position)
-                lane_ids.append(int(lane_id) if lane_id is not None else None)
+                lane_ids.append(str(lane_id) if lane_id is not None else None)
             else:
                 lane_ids.append(None)
         if not len(lane_ids) == self.num_timesteps:
@@ -121,13 +118,14 @@ class ActorGraph:
 
         return lane_ids
 
[…]

### 2025-04-07 graph_creator/ActorGraph.py
adapt to dict

@@ -79,16 +79,12 @@ class TrackData(BaseModel):
 
 
 class ActorGraph:
-    # add opposite direction
-    # use segmeent length for cut off
-    # go over alle scenarios
-    # add relation for actors on same ID
     def __init__(self):
         self.G_map = None
         self.num_timesteps = None
         self.follow_vehicle_steps = None
         self.track_lane_dict = None
-        self.actor_graphs = []
+        self.actor_graphs = {}
 
     def find_lane_id_from_pos(self, position):
         point = Point(position[0], position[1])
@@ -359,7 +355,6 @@ class ActorGraph:
         timestep_delta = int(len(self.track_lane_dict) / number_graphs)
 
         timestep_graphs = {}
-        print(len(self.track_lane_dict))
         for t in tqdm(range(0, len(self.track_lane_dict), timestep_delta)):
             G_t = nx.MultiDiGraph()
 
@@ -379,25 +374,17 @@ class ActorGraph:
             keys = list(self.track_lane_dict.keys())
             for i in range(len(keys) - 1):
                 track_id_A = keys[i]
-                lane_ids_A = self.track_lane_dict[keys[i]]
-                #for track_id_A, lane_ids_A in self.track_lane_dict.items():
+                lane_ids_A = [ None if lane_id is None else int(lane_id) for lane_id in self.track_lane_dict[keys[i]]]
             
                 if lane_ids_A[t] is None:
                     continue
 
-                #for track_id_B, lane_ids_B in self.track_lane_dict.items():
                 for j in range(i + 1, len(keys)):
                     track_id_B = keys[j]
-                    lane_ids_B = self.track_lane_dict[keys[j]]
+                    lane_ids_B = [ None if lane_id is None else int(lane_id) for lane_id in self.track_lane_dict[keys[j]]]
 
                     if lane_ids_B[t] is None:
                         continue
-                    # Skip if we've already processed this pair.
-                    # Should not happen with the above nested loops, but keep it to be on the safe side.
-                    # if str(track_id_A) == str(track_id_B):
-                    #    continue
-                    #if (track_id_B, track_id_A) in G_t.edges():
-                    #    continue
 
                     # Check for "following_lead" and "leading_vehicle"
                     if nx.has_path(G_map.graph, lane_ids_A[t], lane_ids_B[t]):
@@ -493,7 +480,9 @@ class ActorGraph:
[…]

### 2025-04-07 graph_creator/create_graph.py

@@ -115,4 +115,4 @@ if __name__ == "__main__":
     actor_graph = ActorGraph.from_argoverse_scenario(scenario, G_map)
     show_timestep = 1
     plot_scene_at_timestep(scenario, map, timestep=show_timestep, save_path=(repo_root / "map.png"))
-    actor_graph.visualize_actor_graph(timestep=show_timestep, save_path=(repo_root / "actor_graph.png"))
+    actor_graph.visualize_actor_graph(t_idx=show_timestep, comp_idx=0,  save_path=(repo_root / "actor_graph.png"))

### 2025-08-25 carla/create_data_script_synchronous.py
get the graph embeddings started

@@ -47,14 +47,13 @@ def get_t_coordinate(actor, world_map):
     return t
 
 
-#for j in range(7, 20):
+#for sim_run in range(10):
+
 clean_carla(world)
 _ = world.tick()
-
 client.load_world(random.choice(client.get_available_maps()))
 # create the lane map graph and store it to file:
 world_map = world.get_map()
-
 map_g = MapGraph()
 map_g = map_g.create_from_carla_map(world_map)
 dt_specifier = str(datetime.now())
@@ -68,30 +67,212 @@ settings = world.get_settings()
 settings.synchronous_mode = True  # Enable sync mode
 settings.fixed_delta_seconds = 0.05  # Set fixed time step
 world.apply_settings(settings)
+# BEHAVIOR CONTROL OPTIONS
+# These parameters control actor behavior variability to create more diverse traffic scenarios
+# All boolean flags can be set to False to disable specific behavior variations
+# Speed variation parameters
+SPEED_VARIATION_ENABLED = True
+MIN_SPEED_FACTOR = 0.6  # 60% of speed limit
+MAX_SPEED_FACTOR = 1.4  # 140% of speed limit
+# Distance variation parameters 
+DISTANCE_VARIATION_ENABLED = True
+MIN_DISTANCE_FACTOR = 0.5  # Closer following distance
+MAX_DISTANCE_FACTOR = 3.0  # Larger following distance
+# Aggressive/Conservative driving parameters
+AGGRESSIVE_DRIVING_ENABLED = True
+AGGRESSIVE_PROBABILITY = 0.3  # 30% of vehicles drive aggressively
+# Lane change behavior parameters
+LANE_CHANGE_VARIATION_ENABLED = True
+MIN_LANE_CHANGE_DISTANCE = 10.0  # Meters
+MAX_LANE_CHANGE_DISTANCE = 50.0  # Meters
+# Braking behavior parameters
+BRAKING_VARIATION_ENABLED = True
+MIN_BRAKING_DISTANCE = 5.0  # Meters
+MAX_BRAKING_DISTANCE = 15.0  # Meters
+# Random behavior parameters
+RANDOM_SLOW_DOWNS_ENABLED = True
+SLOW_DOWN_PROBABILITY = 0.1  # 10% chance per step for random slowdown
+SLOW_DOWN_DURATION_RANGE = (50, 200)  # Steps to maintain slow speed
+# Vehicle type variation (affects behavior)
[…]

### 2025-08-28 graph_creator/ActorGraph.py
commit after visit

@@ -152,6 +152,10 @@ class ActorGraph:
         max_distance_opposite_veh_m=100,
         max_distance_neighbor_forward_m=50,
         max_distance_neighbor_backward_m=50,
+        max_number_lead_vehicle=3,
+        max_number_neighbor=3,
+        max_number_opposite=3,
+        max_node_distance=3,
         delta_timestep_s=1.0,
     ):
         """
@@ -190,6 +194,10 @@ class ActorGraph:
             max_distance_neighbor_forward_m=max_distance_neighbor_forward_m,
             max_distance_neighbor_backward_m=max_distance_neighbor_backward_m,
             max_distance_opposite_m=max_distance_opposite_veh_m,
+            max_number_lead_vehicle=max_number_lead_vehicle,
+            max_number_neighbor=max_number_neighbor,
+            max_number_opposite=max_number_opposite,
+            max_node_distance=max_node_distance,
             delta_timestep_s=delta_timestep_s,
         )
 
@@ -204,6 +212,10 @@ class ActorGraph:
         max_distance_opposite_veh_m=100,
         max_distance_neighbor_forward_m=50,
         max_distance_neighbor_backward_m=50,
+        max_number_lead_vehicle=3,
+        max_number_neighbor=3,
+        max_number_opposite=3,
+        max_node_distance=3,
     ):
         """
         Create an ActorGraph instance from a CARLA scenario.
@@ -241,6 +253,10 @@ class ActorGraph:
             max_distance_neighbor_forward_m=max_distance_neighbor_forward_m,
             max_distance_neighbor_backward_m=max_distance_neighbor_backward_m,
             max_distance_opposite_m=max_distance_opposite_veh_m,
+            max_number_lead_vehicle=max_number_lead_vehicle,
+            max_number_neighbor=max_number_neighbor,
+            max_number_opposite=max_number_opposite,
+            max_node_distance=max_node_distance,
         )
         instance.actor_components = {}
         # print("instance.actor_graphs.keys(): ", instance.actor_graphs.keys())
@@ -291,6 +307,77 @@ class ActorGraph:
             track_actor_type_dict=track_actor_type_dict,
         )
 
+    def _has_path_within_distance(self, G, source, target, max_distance):
+        """
+        Check if there's a path from source to target with at most max_distance nodes.
+        
+        Args:
+            G: NetworkX graph
+            source: Source node
+            target: Target node
+            max_distance: Maximum number of nodes in the path
+            
+        Returns:
[…]

### 2025-08-28 graph_creator/create_graph.py

@@ -150,7 +150,7 @@ if __name__ == "__main__":
     dataroot = repo_root / "argoverse_data" / "train"
 
     # unique log identifier
-    log_id = "2c9bc3ec-979e-4873-bd19-13ff38c22cdb"
+    log_id = "d25d1aaa-8bb2-4c6b-98a5-fa5aa2ddd2eb"
 
     scenario, map = get_scenario_data(dataroot, log_id)
     G_map = MapGraph.create_from_argoverse_map(map)
@@ -158,7 +158,14 @@ if __name__ == "__main__":
     # visualize_scenario(scenario, map, save_path=(repo_root / "scenario_plot.mp4") )
     # plot_argoverse_map(map, save_path=(repo_root / "map.png"))
 
-    actor_graph = ActorGraph.from_argoverse_scenario(scenario, G_map)
+    actor_graph = ActorGraph.from_argoverse_scenario(
+        scenario, 
+        G_map,
+        max_number_lead_vehicle=1,
+        max_number_neighbor=1,
+        max_number_opposite=1,
+        max_node_distance=3
+    )
     show_timestep = 1.0
     plot_scene_at_timestep(scenario, map, timestep=show_timestep, save_path=(repo_root / "map.png"))
     actor_graph.visualize_actor_graph(t_idx=show_timestep, comp_idx=0,  save_path=(repo_root / "actor_graph.png"))

### 2025-08-28 graph_creator/ActorGraph.py
Problem: actor can be on multiple overlapping lane IDs

@@ -16,29 +16,38 @@ class ActorGraph:
         self.track_lane_dict = None
         self.actor_graphs = {}
 
-    def find_lane_id_from_pos(self, position):
+    def find_lane_ids_from_pos(self, position):
+        """
+        Find all lane IDs that contain the given position.
+        An actor can be on multiple map elements, so we return a list of lane IDs.
+        Most of the time, there will be just a single element.
+        """
         point = Point(position[0], position[1])
+        lane_ids = []
         for lane_id, data in self.G_map.graph.nodes(data=True):
             lane_polygon = data["node_info"].lane_polygon
             if lane_polygon.contains(point):
-                return lane_id
-        return None
+                lane_ids.append(lane_id)
+        return lane_ids if lane_ids else [None]
 
     def find_lane_ids_for_track(self, track):
         """
-        create a list of length of number of timesteps, where each element holds the lane_id of an object at that timestep.
+        create a list of length of number of timesteps, where each element holds a list of lane_ids of an object at that timestep.
+        An actor can be on multiple map elements, so each element is a list of lane IDs.
+        Most of the time, there will be just a single element in each list.
 
-        Missing information is represented by None.
+        Missing information is represented by [None].
         """
         lane_ids = []
         timestep_list = [step.timestep for step in track.object_states]
         for ii in range(self.num_timesteps):
             if ii in timestep_list:
                 position = track.object_states[timestep_list.index(ii)].position
-                lane_id = self.find_lane_id_from_pos(position)
-                lane_ids.append(str(lane_id) if lane_id is not None else None)
+                lane_id_list = self.find_lane_ids_from_pos(position)
+                # Convert lane IDs to strings, handling the case where lane_id_list might contain None
+                lane_ids.append([str(lane_id) if lane_id is not None else None for lane_id in lane_id_list])
             else:
-                lane_ids.append(None)
+                lane_ids.append([None])
         if not len(lane_ids) == self.num_timesteps:
             raise ValueError(f"There are too many lane IDs for track {track.track_id}")
 
@@ -108,11 +117,17 @@ class ActorGraph:
                     state = track.object_states[timestep_list.index(ii)]
                     position = Point(state.position[0], state.position[1], 0.0)
                     xyz_positions.append(position)
+                    lane_id_list = track_lane_dict[track_id][ii]
                     if (
[…]

### 2025-08-28 graph_creator/models.py

@@ -119,8 +119,8 @@ class ActorType(Enum):
     CYCLIST = "CYCLIST"  # maybe delete this
 
 class TrackData(BaseModel):
-    track_lane_dict: Dict[str, List[Optional[str]]] = Field(
-        description="Dictionary mapping track IDs to lists of lane IDs (can be None)"
+    track_lane_dict: Dict[str, List[List[Optional[str]]]] = Field(
+        description="Dictionary mapping track IDs to lists of lists of lane IDs. Each inner list represents lane IDs for a timestep (can contain None)"
     )
     track_s_value_dict: Dict[str, List[float]] = Field(description="Dictionary mapping track IDs to lists of s-values")
     track_xyz_pos_dict: Dict[str, List[Point]] = Field(description="Dictionary mapping track IDs to lists of 3D points")

### 2025-08-28 graph_creator/MapGraph.py
Better visualizataion of double edges in map - that should not exist

@@ -204,39 +204,79 @@ class MapGraph:
             verticalalignment="bottom",
         )
 
-        # move label away from nodes..
-        # Draw edges with different styles based on edge type
+        # Helper function to offset edges orthogonally to their direction
+        def offset_edge_positions(edges, offset_distance=0.5):
+            """Offset edge positions orthogonally to make overlapping edges visible"""
+            offset_pos = pos.copy()
+            
+            for u, v in edges:
+                if u in pos and v in pos:
+                    # Calculate edge direction vector
+                    dx = pos[v][0] - pos[u][0]
+                    dy = pos[v][1] - pos[u][1]
+                    edge_length = np.sqrt(dx**2 + dy**2)
+                    
+                    if edge_length > 0:
+                        # Normalize and rotate 90 degrees to get orthogonal direction
+                        orthogonal_dx = -dy / edge_length
+                        orthogonal_dy = dx / edge_length
+                        
+                        # Apply offset to both nodes
+                        offset_pos[u] = (pos[u][0] + orthogonal_dx * offset_distance, 
+                                       pos[u][1] + orthogonal_dy * offset_distance)
+                        offset_pos[v] = (pos[v][0] + orthogonal_dx * offset_distance, 
+                                       pos[v][1] + orthogonal_dy * offset_distance)
+            
+            return offset_pos
+
+        # Separate edges by type
         edge_type_fol = [(u, v) for u, v, d in self.graph.edges(data=True) if d["edge_type"] == "following"]
         edge_type_nei = [(u, v) for u, v, d in self.graph.edges(data=True) if d["edge_type"] == "neighbor"]
         edge_type_opp = [(u, v) for u, v, d in self.graph.edges(data=True) if d["edge_type"] == "opposite"]
 
+        # Draw edges with offsets to prevent overlap
+        fol_pos = offset_edge_positions(edge_type_fol, offset_distance=0.3)
+        nei_pos = offset_edge_positions(edge_type_nei, offset_distance=-0.3)
+        opp_pos = offset_edge_positions(edge_type_opp, offset_distance=0.6)
+
+        # Draw following edges (blue)
         nx.draw_networkx_edges(
             self.graph,
-            pos,
+            fol_pos,
             edgelist=edge_type_fol,
             width=2,
             edge_color="blue",
             node_size=node_size,
+            label="Following",
         )
+        
+        # Draw neighbor edges (green)
         nx.draw_networkx_edges(
             self.graph,
-            pos,
[…]

### 2025-08-28 graph_creator/MapGraph.py
bug in argoverse MapGraph

@@ -37,12 +37,18 @@ class MapGraph:
         for lane_id, lane in map.vector_lane_segments.items():
             if lane.left_neighbor_id is not None:
                 if str(lane.left_neighbor_id) in G:
-                    G.add_edge(str(lane_id), str(lane.left_neighbor_id), edge_type="neighbor")
-                    G.add_edge(str(lane.left_neighbor_id), str(lane_id), edge_type="neighbor")
+                    # Only add if the edge doesn't already exist
+                    if not G.has_edge(str(lane_id), str(lane.left_neighbor_id)):
+                        G.add_edge(str(lane_id), str(lane.left_neighbor_id), edge_type="neighbor")
+                    if not G.has_edge(str(lane.left_neighbor_id), str(lane_id)):
+                        G.add_edge(str(lane.left_neighbor_id), str(lane_id), edge_type="neighbor")
             if lane.right_neighbor_id is not None:
                 if str(lane.right_neighbor_id) in G:
-                    G.add_edge(str(lane_id), str(lane.right_neighbor_id), edge_type="neighbor")
-                    G.add_edge(str(lane.right_neighbor_id), str(lane_id), edge_type="neighbor")
+                    # Only add if the edge doesn't already exist
+                    if not G.has_edge(str(lane_id), str(lane.right_neighbor_id)):
+                        G.add_edge(str(lane_id), str(lane.right_neighbor_id), edge_type="neighbor")
+                    if not G.has_edge(str(lane.right_neighbor_id), str(lane_id)):
+                        G.add_edge(str(lane.right_neighbor_id), str(lane_id), edge_type="neighbor")
 
         # Rename neighboring lanes from lanes in opposite direction by looking for loops.
         edges_opposite = []
@@ -57,10 +63,12 @@ class MapGraph:
                                         edges_opposite.append((successor, neighbor))
                                         edges_opposite.append((next_node, node))
 
-        # Rename edges to 'opposite'
+        # Convert neighbor edges to opposite edges
         for u, v in edges_opposite:
+            # Just rename the edge type from 'neighbor' to 'opposite'
             G[u][v][0]["edge_type"] = "opposite"
-
+            G[v][u][0]["edge_type"] = "opposite"
+        
         return instance
 
     def _to_2d(self, location):

### 2025-08-30 graph_creator/ActorGraph.py
graph embeddings

@@ -496,8 +496,8 @@ class ActorGraph:
 
                         # TODO: PROBLEM: we are not checking i in relationt j,but not j to 1. We might miss a case here
 
-                        if track_id_A == "188747" and track_id_B == "188849":
-                            print(f"path: {path}")
+                        # if track_id_A == "188747" and track_id_B == "188849":
+                        #     print(f"path: {path}")
                         # third case: on neighboring lanes, forward
                         if (
                             sum([G_map.graph[u][v][0]["edge_type"] == "neighbor" for u, v in zip(path[:-1], path[1:])])
@@ -653,8 +653,8 @@ class ActorGraph:
                         lane_change = False
                     else:
                         lane_change = True
-                    if lane_change:
-                        print(node, previous_lane_id, lane_id, start_points, lane_change)
+                    #if lane_change:
+                    #     # print(node, previous_lane_id, lane_id, start_points, lane_change)
                 # In principle, here there could also be added a check for lane merge, i.e. counting if the start points had more then 1 element (before adding the current lane_id)
                 else:
                     lane_change = False

### 2025-09-13 graph_creator/ActorGraph.py
redo embeddings with new argoverse mapgraphs, do a subgraph isomorphism nb

@@ -237,6 +237,7 @@ class ActorGraph:
         max_node_distance_leading=3,
         max_node_distance_neighbor=3,
         max_node_distance_opposite=3,
+        delta_timestep_s=1.0,
     ):
         """
         Create an ActorGraph instance from a CARLA scenario.
@@ -283,6 +284,7 @@ class ActorGraph:
             max_node_distance_leading=max_node_distance_leading,
             max_node_distance_neighbor=max_node_distance_neighbor,
             max_node_distance_opposite=max_node_distance_opposite,
+            delta_timestep_s=delta_timestep_s,
         )
         instance.actor_components = {}
         # print("instance.actor_graphs.keys(): ", instance.actor_graphs.keys())

### 2025-09-13 article/article.tex
try out graph staff and add references
some start with paper writing

@@
 
-\title{Using graphs for advanced coverage arguments in autonomous driving.}
+\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}
 \author{xxx}
@@
 
-\section{exsting coverage approaches}
+\begin{itemize}
+    \item In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
+    \item In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
+    \item See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
+    \item In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
+    \item There are already other graph-based approaches for analysing and representing traffic scenes, see for example \cite{behr2021graph}.
+    \item However, the work in that paper is not specifically focused on coverage analysis.
+    \item Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
+    \item This paper is structured as follows:
+    \begin{itemize}
+        \item In the first section, xxx
+    \end{itemize}
+\end{itemize}
+
+\cite{wagner2022odd}
+
+\cite{arzamasov2021data}
+
+\section{exsting coverage and analysis approaches}
+
+add Master thesis Johannes
+
+\cite{pegasus2019method}
+
+\cite{Ries2021traj_clustering}
+
+\cite{ORAD2021taxonomy}
+
+\cite{Ulbrich2015scene}
+
+\cite{DBLP:journals/corr/abs-1801-08598}
+
+\cite{Ammann_Offutt_2008}
+
+\cite{foretellix2019}
+
+\cite{deGelder2022ontology}
+
+\cite{wachenfeld2016release}
+
+\cite{berger2020survey}
+
+\cite{iso21448}
+
+\cite{ul4600}
 
@@
 
-\section{Analysing a traffic scene graph from a coverage perspective}
+\cite{behr2021graph}
+
 \cite{newman2010networks}
+
+\cite{bagheri2020ontology}
+
+\cite{riedmaier2020realistic}
+
+\cite{fremont2020scenic}
+
+\cite{pek2019generating}
+
+\subsection{time based graph representations}
+
+\section{Analysing a traffic scene with agraph}
+
+\begin{itemize}
+    \item Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
+    \item Two methodologies are proposed for this purpose:
+    \item One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
+    \item The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
+\end{itemize}
+
+\subsection{Create subgraphs for coverage analysis}
+
+\begin{itemize}
+    \item There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
+    \item Once an archetype is defined, a special property of graphs can be used. 
+    \item Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
+    \item As the archetypes are not necessarily are involving a lot of actors, these are more like subsets of actual traffic scenes.
+    \item A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
+    \item This situation can be represented by a graph with 3 nodes and 2 edges. 
+    \item In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
+    \item This is an example of a subgraph isomorphism problem.
+    \item While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
+    \item One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library. <- Check and cite something.
+    \item The strategy we are then applying is 
+\end{itemize}
+
+\subsection{graph embeddings for coverage analysis}
 
@@
 
+\cite{Argoverse2}
+
 \subsection{Carla}
 
-\cite{goodfellow2016deep}
+\cite{Dosovitskiy17Carla}
+
+% \cite{goodfellow2016deep} % needed?
 

### 2025-09-14 carla/create_data_script_synchronous.py
correct actor graph

@@ -5,7 +5,7 @@
 import pandas as pd
 import os
 
-for i in range(10):
+for i in range(30):
     os.chdir("carla")
     from src.generate_traffic_data import clean_carla, spawn_scene, run_scene
     import carla

### 2025-09-14 carla/create_data_script_synchronous.py
some small updates

@@ -5,7 +5,7 @@
 import pandas as pd
 import os
 
-for i in range(30):
+for ijk in range(30):
     os.chdir("carla")
     from src.generate_traffic_data import clean_carla, spawn_scene, run_scene
     import carla
@@ -251,7 +251,7 @@ for i in range(30):
                             speed_percentage = (speed_factor - 1.0) * 100.0
                             tm.vehicle_percentage_speed_difference(act, speed_percentage)        
         # Apply periodic behavior variations every 500 steps
-        if i % 500 == 0 and i > 0:
+        if i % 100 == 0 and i > 0:
             for act in world.get_actors().filter("vehicle.*"):
                 # Randomly change some vehicle behaviors
                 if random.random() < 0.2:  # 20% chance to change behavior

### 2025-09-14 article/article.tex

@@
     \item One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library. <- Check and cite something.
-    \item The strategy we are then applying is 
+    \item The strategy we are then applying is a simple counting strategy.
+    \item TODO: Create pseudo code for double loop counting strategy.
+    \item This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
+    \item Then going upwards to different datasets, it is checked, if the archetype is present.
 \end{itemize}
@@
 \subsection{graph embeddings for coverage analysis}
+
+\begin{itemize}
+    \item One often utilized strategy in machine learning is to translate raw data like images or text into an embedding space in order to be  
+\end{itemize}
 

### 2025-09-24 article/article.tex
a little bit of writing, a little bit of subgraph isomorphism

@@
     \item While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-    \item One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library. <- Check and cite something.
-    \item The strategy we are then applying is a simple counting strategy.
-    \item TODO: Create pseudo code for double loop counting strategy.
+    \item One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
+    \item The strategy we are then applying is the following:
+    \item \begin{enumerate}
+            \item Define a set of subgraphs $S$ that are considered to be archetypes.
+            \item Define which node and edge attributes are considered for the isomorphism check.
+            \item Create an empty dataframe $C$ with a column for each subgraph in $S$
+            \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
+            \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
+        \end{enumerate}
     \item This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
     \item Then going upwards to different datasets, it is checked, if the archetype is present.
+    \item Also, follow up analysis of the created coverage dataframe can be performed. For example,
+    \item \begin{itemize}
+        \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
+        \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
+        \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
+    \end{itemize}
     \end{itemize}

### 2025-09-24 graph_creator/utilities.py

@@ -1,0 +1,21 @@
+import pandas as pd
+import networkx as nx
+from typing import Tuple
+
+def make_node_edge_df(graph: nx.Graph) -> Tuple[pd.DataFrame, pd.DataFrame]:
+    """
+    Make a dataframe from a networkx graph.
+    
+    Args:
+        graph: networkx graph
+
+    Returns:
+        df_nodes: dataframe of nodes
+        df_edges: dataframe of edges
+    """
+    df_nodes = pd.DataFrame.from_dict(dict(graph.nodes(data=True)), orient='index')
+    df_nodes.reset_index(inplace=True)
+    df_nodes.rename(columns={'index': 'node_id'}, inplace=True)
+    df_edges = pd.DataFrame([(u, v, d) for u, v, d in graph.edges(data=True)], columns=['source', 'target', 'attributes'])
+    df_edges = pd.concat([df_edges[['source', 'target']], pd.json_normalize(df_edges['attributes'])], axis=1)
+    return df_nodes, df_edges

### 2025-09-28 article/article.tex
small additions to article
small part of paper writing

@@
 
+\section{Implementation of Graph Embeddings for Traffic Scene Analysis}
+
+This section describes the implementation and application of graph embeddings to traffic scene graphs, as demonstrated in the \texttt{graph\_embeddings.ipynb} notebook.
+The implementation follows a comprehensive approach to learning graph representations through self-supervised contrastive learning.
+
+\subsection{Graph Neural Network Architecture}
+
+The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.
+The model consists of multiple GINE convolutional layers, each containing:
+\begin{itemize}
+    \item Sequential multi-layer perceptrons (MLPs) with batch normalization and ReLU activations
+    \item Edge-aware message passing incorporating both node and edge features
+    \item Dropout regularization to prevent overfitting
+\end{itemize}
+
+The model employs a combination of three graph pooling strategies - mean, max, and sum pooling - to create a comprehensive graph-level representation.
+A projection head enables contrastive learning through InfoNCE loss, while an optional classification head supports supervised learning tasks.
+
+\subsection{Data Processing and Augmentation}
+
+The \texttt{GraphDataset} class handles the conversion from NetworkX graphs to PyTorch Geometric format through the \texttt{networkx\_to\_pyg} function. This conversion includes:
+\begin{itemize}
+    \item One-hot encoding of categorical node features (actor types: VEHICLE, PEDESTRIAN, CYCLIST, MOTORCYCLE)
+    \item One-hot encoding of edge types (neighbor\_vehicle, opposite\_vehicle, same\_lane, adjacent\_lane, following, intersection)
+    \item Preservation of continuous features such as longitudinal speed and path length
+\end{itemize}
+
+Data augmentation is implemented through the \texttt{augment\_graph} function, which adds Gaussian noise to continuous features to create augmented versions for contrastive learning.
+
+\subsection{Training Methodology}
+
+The training process utilizes self-supervised contrastive learning with the following characteristics:
+\begin{itemize}
+    \item InfoNCE contrastive loss implemented in the \texttt{contrastive\_loss} function
+    \item Adaptive learning rate schedule with exponential decay (starting at 0.02, decaying by factor 0.75)
+    \item Multi-epoch training with 15 epochs per learning rate stage
+    \item Data from both CARLA simulator (4,442 graphs) and Argoverse 2.0 dataset (3,588 graphs)
+\end{itemize}
+
+\subsection{Embedding Analysis and Visualization}
+
+The trained model generates 256-dimensional embeddings for each traffic scene graph. Analysis includes:
+\begin{itemize}
+    \item Dimensionality reduction using Principal Component Analysis (PCA) and t-distributed Stochastic Neighbor Embedding (t-SNE)
+    \item Visualization of embedding space distinguishing between CARLA and Argoverse data sources
+    \item Similarity-based scene retrieval using squared Euclidean distance in embedding space
+    \item Integration with visualization functions (\texttt{plot\_lane\_map\_advanced}, \texttt{add\_actors\_to\_map}, \texttt{add\_actor\_edges\_to\_map}) to display similar traffic scenes
+\end{itemize}
+
+The implementation demonstrates successful learning of traffic scene representations, with the embedding space capturing meaningful similarities between scenes from different data sources while maintaining distinction between simulation and real-world data.
+
 \section{Application}

### 2025-09-30 carla/create_data_script_synchronous.py
some debugging of Carla data

@@ -5,16 +5,13 @@
 import pandas as pd
 import os
 os.chdir("carla")
-from src.generate_traffic_data import clean_carla# , spawn_scene, run_scene
+from src.generate_traffic_data import clean_carla
 import carla
 os.getcwd()
 from datetime import datetime
 import time
 import random
 from tqdm import tqdm
-#import networkx as nx
-#import numpy as np
-#from shapely.geometry import Polygon
 os.chdir("..")
 os.getcwd()
 from graph_creator.MapGraph import MapGraph

### 2025-09-30 carla/on_or_offroad.py

@@ -1,0 +1,50 @@
+# This script is used to check if an actor is on a road or offroad and to store the result to the tracks parquet file
+# precisely, the smallest distance to any lane in the map graph is calculated and stored in the tracks parquet file
+
+import pandas as pd
+import os
+#os.chdir("carla")
+from graph_creator.MapGraph import MapGraph
+from shapely.geometry import Point
+from tqdm import tqdm
+import numpy as np
+
+files = os.listdir("/home/tmuehlen/repos/graph_coverage/carla/data")
+scn_ids = [file.split("_")[1] for file in files if "tracks" in file]
+scn_ids = sorted(scn_ids)
+scn_ids = [scn_id for scn_id in scn_ids if scn_id > "2025-09-05"]
+print(len(scn_ids))
+
+#scn_id = scn_ids[100]
+for j, scn_id in enumerate(scn_ids):
+    print(f'{j}/{len(scn_ids)}: {scn_id}')
+    tracks = pd.read_parquet(f'/home/tmuehlen/repos/graph_coverage/carla/data/scene_{scn_id}_tracks.parquet')
+    g_map = MapGraph()
+    g_map.read_graph_from_file(f'/home/tmuehlen/repos/graph_coverage/carla/data/scene_{scn_id}_map_graph.pickle')
+    distances_orig_lane = []
+    distances_min_lane = []
+    min_distances_lane_ids = []
+    for i in tqdm(range(len(tracks))):
+        road_lane_id = str(tracks.road_id.iloc[i]) + '_' + str(tracks.lane_id.iloc[i])
+        distance_orig_lane = Point(tracks.actor_location_xyz.iloc[i]).distance(g_map.graph.nodes[road_lane_id]["node_info"].lane_polygon)
+        if distance_orig_lane > 0.0:
+            min_distance = np.inf
+            for lane in g_map.graph.nodes(data=True):
+                distance = Point(tracks.actor_location_xyz.iloc[i]).distance(lane[1]["node_info"].lane_polygon)
+                if distance < min_distance:
+                    min_distance = distance
+                    min_distance_lane_id = lane[0]
+            distances_orig_lane.append(distance_orig_lane)
+            distances_min_lane.append(min_distance)
+            min_distances_lane_ids.append(min_distance_lane_id)
+            # print("min distance details: ", i,  distance_orig_lane, min_distance, road_lane_id, min_distance_lane_id)
+        else:
[…]

### 2025-10-02 carla/on_or_offroad.py
remove offroad actors from graph

@@ -10,7 +10,7 @@ from tqdm import tqdm
 import numpy as np
 
 files = os.listdir("/home/tmuehlen/repos/graph_coverage/carla/data")
-scn_ids = [file.split("_")[1] for file in files if "tracks" in file]
+scn_ids = [file.split("_")[1] for file in files if "tracks.parquet" in file]
 scn_ids = sorted(scn_ids)
 scn_ids = [scn_id for scn_id in scn_ids if scn_id > "2025-09-05"]
 print(len(scn_ids))
@@ -30,10 +30,13 @@ for j, scn_id in enumerate(scn_ids):
         if distance_orig_lane > 0.0:
             min_distance = np.inf
             for lane in g_map.graph.nodes(data=True):
-                distance = Point(tracks.actor_location_xyz.iloc[i]).distance(lane[1]["node_info"].lane_polygon)
-                if distance < min_distance:
-                    min_distance = distance
-                    min_distance_lane_id = lane[0]
+                try:
+                    distance = Point(tracks.actor_location_xyz.iloc[i]).distance(lane[1]["node_info"].lane_polygon)
+                    if distance < min_distance:
+                        min_distance = distance
+                        min_distance_lane_id = lane[0]
+                except:
+                    print("error", i, road_lane_id)
             distances_orig_lane.append(distance_orig_lane)
             distances_min_lane.append(min_distance)
             min_distances_lane_ids.append(min_distance_lane_id)

### 2025-10-02 carla/create_data_script_synchronous_simple.py
update colors in map graphs, carla debugging

@@ -1,0 +1,218 @@
+# Before running this script, make sure to start the CARLA server via nomachine (needs a screen):
+# tmuehlen@tmuehlen-HP-Z8-G4-Workstation:~/repos/graph_coverage/CARLA_0.9.15$ ./CarlaUE4.sh 
+# use `poetry shell` to activate the environment
+
+import pandas as pd
+import os
+os.chdir("carla")
+from src.generate_traffic_data import clean_carla
+import carla
+os.getcwd()
+from datetime import datetime
+import time
+import random
+from tqdm import tqdm
+os.chdir("..")
+os.getcwd()
+from graph_creator.MapGraph import MapGraph
+
+def get_t_coordinate(actor, world_map):
+    # Thanks Gemini
+    location = actor.get_location()
+    # Get the closest waypoint on a driving lane
+    waypoint = world_map.get_waypoint(location, project_to_road=True, lane_type=carla.LaneType.Driving)
+    # Center of the lane
+    lane_center = waypoint.transform.location
+    # Right vector of the lane (perpendicular to the forward direction)
+    right_vector = waypoint.transform.get_right_vector()
+    # Vector from lane center to actor
+    offset_vector = location - lane_center
+    # Compute t-coordinate: lateral distance (positive to the right of lane center)
+    t = right_vector.x * offset_vector.x + right_vector.y * offset_vector.y + right_vector.z * offset_vector.z
+    return t
+
+def to_2d(location):
+    return (location.x, location.y)
+
+    
+for ijk in range(30):
+    client = carla.Client("localhost", 2000)
+    client.set_timeout(50.0)
+    world = client.get_world()
+    clean_carla(world)
+    _ = world.tick()
+    client.get_available_maps()
+    maps = [
+        '/Game/Carla/Maps/Town01', 
+        '/Game/Carla/Maps/Town01_Opt', 
+        '/Game/Carla/Maps/Town02_Opt', 
+        '/Game/Carla/Maps/Town02', 
+        '/Game/Carla/Maps/Town03', 
+        '/Game/Carla/Maps/Town04', 
+        '/Game/Carla/Maps/Town04_Opt', 
+        '/Game/Carla/Maps/Town05_Opt', 
+        '/Game/Carla/Maps/Town05', 
+        '/Game/Carla/Maps/Town07', 
+        '/Game/Carla/Maps/Town10HD', 
+        '/Game/Carla/Maps/Town10HD_Opt', 
[…]

### 2025-10-02 graph_creator/plot_graphs.py

@@ -261,10 +261,8 @@ def add_actor_edges_to_map(fig, ax, actor_graph):
    # Color mapping for different edge types
    edge_colors = {
        'opposite_vehicle': 'red',
-       'same_lane': 'blue', 
-       'adjacent_lane': 'green',
-       'intersection': 'orange',
-       'following': 'purple'
+       'following_lead': 'purple',
+       'neighbor_vehicle': 'green',
    }
    default_edge_color = 'gray'
    
@@ -286,8 +284,8 @@ def add_actor_edges_to_map(fig, ax, actor_graph):
        
        # Line style based on path length (negative = dashed, positive = solid)
        linestyle = '--' if path_length < 0 else '-'
-       linewidth = min(max(abs(path_length) / 10, 0.5), 3.0)  # Scale line width by path length
-       
+       # linewidth = min(max(abs(path_length) / 10, 0.5), 3.0)  # Scale line width by path length
+       linewidth = 0.75
        # Draw edge as arrow
        ax.annotate('', xy=(target_x, target_y), xytext=(source_x, source_y),
                   arrowprops=dict(arrowstyle='->', 

### 2025-10-02 article/article.tex

@@
 
+CARLA (Car Learning to Act) is an open-source simulator specifically designed for autonomous driving research and development.
+It provides a highly realistic urban driving environment with diverse road layouts, weather conditions, and traffic scenarios.
+The simulator features a comprehensive sensor suite simulation, flexible API for scenario creation, and supports both learning-based and traditional autonomous driving approaches.
+CARLA enables researchers to test and validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
+
+The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
+benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}.
+Industry applications include
+virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}.
+CARLA is also extensively used in autonomous driving competitions and challenges, providing a common evaluation environment for comparing different approaches across research groups worldwide.
+
 \cite{Dosovitskiy17Carla}

### 2025-10-05 carla/create_data_script_synchronous.py
debug carla map error

@@ -34,346 +34,350 @@ def get_t_coordinate(actor, world_map):
 def to_2d(location):
     return (location.x, location.y)
 
-    
-for ijk in range(30):
-    client = carla.Client("localhost", 2000)
-    client.set_timeout(50.0)
-    world = client.get_world()
-    clean_carla(world)
-    _ = world.tick()
-    client.get_available_maps()
-    maps = [
-        '/Game/Carla/Maps/Town01', 
-        '/Game/Carla/Maps/Town01_Opt', 
-        '/Game/Carla/Maps/Town02_Opt', 
-        '/Game/Carla/Maps/Town02', 
-        '/Game/Carla/Maps/Town03', 
-        '/Game/Carla/Maps/Town04', 
-        '/Game/Carla/Maps/Town04_Opt', 
-        '/Game/Carla/Maps/Town05_Opt', 
-        '/Game/Carla/Maps/Town05', 
-        '/Game/Carla/Maps/Town07', 
-        '/Game/Carla/Maps/Town10HD', 
-        '/Game/Carla/Maps/Town10HD_Opt', 
-        '/Game/Carla/Maps/Town11/Town11' 
-    ]
-    map = random.choice(maps)
-    print(map)
-    client.load_world(map)
-    # create the lane map graph and store it to file:
-    world_map = world.get_map()
-    map_g = MapGraph()
-    map_g = map_g.create_from_carla_map(world_map)
-    dt_specifier = str(datetime.now())
-    map_g.store_graph_to_file(f"carla/data/scene_{dt_specifier}_map_graph.pickle")
-    xodr = world_map.to_opendrive()
-    with open(f"carla/data/scene_{dt_specifier}_map.xodr", "w") as f:
-        f.write(xodr)
-    # Get the world and set synchronous mode
-    settings = world.get_settings()
-    settings.synchronous_mode = True  # Enable sync mode
-    settings.fixed_delta_seconds = 0.05  # Set fixed time step
-    world.apply_settings(settings)
-    # BEHAVIOR CONTROL OPTIONS
-    # These parameters control actor behavior variability to create more diverse traffic scenarios
-    # All boolean flags can be set to False to disable specific behavior variations
-    # Speed variation parameters
-    SPEED_VARIATION_ENABLED = True
-    MIN_SPEED_FACTOR = 0.6  # 60% of speed limit
-    MAX_SPEED_FACTOR = 1.4  # 140% of speed limit
-    # Distance variation parameters 
[…]

### 2025-10-05 carla/create_data_script_synchronous_simple.py

@@ -16,6 +16,7 @@ os.chdir("..")
 os.getcwd()
 from graph_creator.MapGraph import MapGraph
 
+
 def get_t_coordinate(actor, world_map):
     # Thanks Gemini
     location = actor.get_location()
@@ -31,188 +32,188 @@ def get_t_coordinate(actor, world_map):
     t = right_vector.x * offset_vector.x + right_vector.y * offset_vector.y + right_vector.z * offset_vector.z
     return t
 
+
 def to_2d(location):
     return (location.x, location.y)
 
-    
-for ijk in range(30):
-    client = carla.Client("localhost", 2000)
-    client.set_timeout(50.0)
-    world = client.get_world()
-    clean_carla(world)
-    _ = world.tick()
-    client.get_available_maps()
-    maps = [
-        '/Game/Carla/Maps/Town01', 
-        '/Game/Carla/Maps/Town01_Opt', 
-        '/Game/Carla/Maps/Town02_Opt', 
-        '/Game/Carla/Maps/Town02', 
-        '/Game/Carla/Maps/Town03', 
-        '/Game/Carla/Maps/Town04', 
-        '/Game/Carla/Maps/Town04_Opt', 
-        '/Game/Carla/Maps/Town05_Opt', 
-        '/Game/Carla/Maps/Town05', 
-        '/Game/Carla/Maps/Town07', 
-        '/Game/Carla/Maps/Town10HD', 
-        '/Game/Carla/Maps/Town10HD_Opt', 
-        '/Game/Carla/Maps/Town11/Town11' 
-    ]
-    map = random.choice(maps)
-    print(map)
-    client.load_world(map)
-    # create the lane map graph and store it to file:
-    world_map = world.get_map()
-    map_g = MapGraph()
-    map_g = map_g.create_from_carla_map(world_map)
-    dt_specifier = str(datetime.now())
-    map_g.store_graph_to_file(f"carla/data/scene_{dt_specifier}_map_graph.pickle")
-    xodr = world_map.to_opendrive()
-    with open(f"carla/data/scene_{dt_specifier}_map.xodr", "w") as f:
-        f.write(xodr)
-    # Get the world and set synchronous mode
-    settings = world.get_settings()
-    settings.synchronous_mode = True  # Enable sync mode
-    settings.fixed_delta_seconds = 0.05  # Set fixed time step
-    world.apply_settings(settings)
 
-    # Get the traffic manager
-    tm = client.get_trafficmanager(8000)  # Port 8000
[…]

### 2025-10-05 graph_creator/models.py

@@ -78,44 +78,56 @@ class NodeInfo(BaseModel):
 
     @classmethod
     def from_carla_lane(cls, lane) -> 'NodeInfo':
-        """Create a NodeInfo instance from a Carla lane object.
-        
-        Args:
-            lane: A Carla lane object containing lane information
-            
-        Returns:
-            NodeInfo instance with the lane information
-        """
+        """Create a NodeInfo instance from a Carla lane object."""
         import carla
-        # TODO @Thomas
+        
+        # Get waypoints along the lane for better polygon accuracy
         wps = lane[0].next_until_lane_end(0.25)
-        lane_length = sum(
-            [wps[i].transform.location.distance(wps[i + 1].transform.location) for i in range(len(wps) - 1)]
-        )
-        # Create lane polygon
-        forward_vector = lane[0].transform.get_forward_vector()
-        right_vector = carla.Vector3D(-forward_vector.y, forward_vector.x, 0)  # Perpendicular to forward
-        # Compute lane width
-        lane_width = lane[0].lane_width
-        # Compute boundary points
-        left_boundary_start = lane[0].transform.location + right_vector * (lane_width / 2.0)
-        right_boundary_start = lane[0].transform.location - right_vector * (lane_width / 2.0)
-        left_boundary_end = lane[1].transform.location + right_vector * (lane_width / 2.0)
-        right_boundary_end = lane[1].transform.location - right_vector * (lane_width / 2.0)
-        lane_polygon = Polygon(
-            [
-                _to_2d(left_boundary_start),
-                _to_2d(left_boundary_end),
-                _to_2d(right_boundary_end),
-                _to_2d(right_boundary_start),
-            ]
+        if len(wps) < 2:
+            wps = [lane[0], lane[1]]  # fallback for very short segments
+        
+        lane_length = sum([
+            wps[i].transform.location.distance(wps[i + 1].transform.location) 
+            for i in range(len(wps) - 1)
+        ])
+        
+        # Create left and right boundary points separately
+        left_boundary_points = []
+        right_boundary_points = []
+        
+        for wp in wps:
+            forward_vector = wp.transform.get_forward_vector()
+            right_vector = carla.Vector3D(-forward_vector.y, forward_vector.x, 0)
+            lane_width = wp.lane_width
+            
+            left_point = wp.transform.location + right_vector * (lane_width / 2.0)
+            right_point = wp.transform.location - right_vector * (lane_width / 2.0)
+            
[…]

### 2025-10-05 article/article.tex
some parts in paper writing

@@
 \usepackage{hyperref}
+\usepackage{graphicx}
 
@@
 
-CARLA (Car Learning to Act) is an open-source simulator specifically designed for autonomous driving research and development.
-It provides a highly realistic urban driving environment with diverse road layouts, weather conditions, and traffic scenarios.
-The simulator features a comprehensive sensor suite simulation, flexible API for scenario creation, and supports both learning-based and traditional autonomous driving approaches.
-CARLA enables researchers to test and validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
+CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
+development. It provides a highly realistic urban driving environment with 
+diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
+sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
+and traditional autonomous driving approaches. CARLA enables researchers to test and 
+validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
 
 The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}.
-Industry applications include
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}.
-CARLA is also extensively used in autonomous driving competitions and challenges, providing a common evaluation environment for comparing different approaches across research groups worldwide.
+benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
+techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
+virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
+with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
+is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
+environment for comparing different approaches across research groups worldwide.
 
-\cite{Dosovitskiy17Carla}
+Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
+Mine\_1 (which is not really normal roads) at the start of this project.
+Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
+of these maps are shown in Figure \ref{fig:carla_maps}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
+\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
+These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
+\label{fig:carla_maps}
+\end{figure}
+
+The data generation script implements sophisticated behavior control mechanisms to create 
+diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
+and regular cars are spawned with varying probabilities, each exhibiting different behavioral
+characteristics such as speed preferences, following distances, and lane-changing tendencies. 
+The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
+periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
+varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
+have between 20 and 60 vehicles each.
+
+The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
+similar data size as the Argoverse 2.0 dataset.
 

### 2025-10-05 carla/create_data_script_synchronous.py

@@ -37,15 +37,15 @@ def to_2d(location):
 
 maps = [
     '/Game/Carla/Maps/Town01', 
-    '/Game/Carla/Maps/Town01_Opt', 
-    '/Game/Carla/Maps/Town02_Opt', 
     '/Game/Carla/Maps/Town02', 
     '/Game/Carla/Maps/Town03', 
     '/Game/Carla/Maps/Town04', # has a lot of offroad actors
-    '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors
-    '/Game/Carla/Maps/Town05_Opt', 
     '/Game/Carla/Maps/Town05', 
     '/Game/Carla/Maps/Town07', 
+    # '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors
+    # '/Game/Carla/Maps/Town05_Opt', 
+    # '/Game/Carla/Maps/Town01_Opt', 
+    # '/Game/Carla/Maps/Town02_Opt', 
     # '/Game/Carla/Maps/Town10HD', # problematic. for whatever reason...
     # '/Game/Carla/Maps/Town10HD_Opt', 
     # '/Game/Carla/Maps/Town11/Town11' 

### 2025-10-05 carla/create_data_script_synchronous_simple.py

@@ -38,16 +38,16 @@ def to_2d(location):
 
 
 maps = [
-    '/Game/Carla/Maps/Town01', 
-    '/Game/Carla/Maps/Town01_Opt', 
-    '/Game/Carla/Maps/Town02_Opt', 
-    '/Game/Carla/Maps/Town02', 
-    '/Game/Carla/Maps/Town03', 
-    '/Game/Carla/Maps/Town04', # has a lot of offroad actors
-    '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors
-    '/Game/Carla/Maps/Town05_Opt', 
+    # '/Game/Carla/Maps/Town01', 
+    # '/Game/Carla/Maps/Town02', 
+    # '/Game/Carla/Maps/Town03', 
+    # '/Game/Carla/Maps/Town04', 
     '/Game/Carla/Maps/Town05', 
     '/Game/Carla/Maps/Town07', 
+    #'/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors
+   # '/Game/Carla/Maps/Town05_Opt', 
+    #'/Game/Carla/Maps/Town01_Opt', 
+    # '/Game/Carla/Maps/Town02_Opt', 
     # '/Game/Carla/Maps/Town10HD', # problematic. for whatever reason...
     # '/Game/Carla/Maps/Town10HD_Opt', 
     # '/Game/Carla/Maps/Town11/Town11' 

### 2025-10-05 article/article.tex
start writing embedding chapter.

@@
 
-\subsection{Create subgraphs for coverage analysis}
+\section{Create subgraphs for coverage analysis}
 
@@
 
-\subsection{graph embeddings for coverage analysis}
-
-\begin{itemize}
-    \item One often utilized strategy in machine learning is to translate raw data like images or text into an embedding space in order to be  
-\end{itemize}
-
 \section{Implementation of Graph Embeddings for Traffic Scene Analysis}
 
-This section describes the implementation and application of graph embeddings to traffic scene graphs, as demonstrated in the \texttt{graph\_embeddings.ipynb} notebook.
-The implementation follows a comprehensive approach to learning graph representations through self-supervised contrastive learning.
+This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
+The implementation follows a comprehensive approach to learning graph representations through self-supervised 
+contrastive learning.
 
-\subsection{Graph Neural Network Architecture}
+Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
+able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
+translate words into a 
+vector space, where the distance between vectors can be used to measure the similarity between 
+words (\cite{mikolov2013efficientestimationwordrepresentations}).
+
+In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
+where the distance between vectors can be used to measure the similarity between traffic scenes.
+This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
+For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
+to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
+to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
+would not be possible.
+
+Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
+gained a lot of popularity in the last years, see for example (add references).
+
+In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
+in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
+pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
+learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
+network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
 

### 2025-10-05 carla/create_data_script_synchronous.py

@@ -36,10 +36,10 @@ def to_2d(location):
 
 
 maps = [
-    '/Game/Carla/Maps/Town01', 
-    '/Game/Carla/Maps/Town02', 
-    '/Game/Carla/Maps/Town03', 
-    '/Game/Carla/Maps/Town04', # has a lot of offroad actors
+    # '/Game/Carla/Maps/Town01', 
+    # '/Game/Carla/Maps/Town02', 
+    # '/Game/Carla/Maps/Town03', 
+    # '/Game/Carla/Maps/Town04', # has a lot of offroad actors
     '/Game/Carla/Maps/Town05', 
     '/Game/Carla/Maps/Town07', 
     # '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors

### 2025-10-06 article/article.tex
work on the embeddings

@@
 
+The exact architecture of the model is shown in Figure \ref{fig:model_architecture}. The features used are the actor type (as a one hot encoding),
+the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its lane since the last timestep (boolean) for the nodes.
+For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
+
 The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.

### 2025-10-06 carla/create_data_script_synchronous.py

@@ -36,10 +36,10 @@ def to_2d(location):
 
 
 maps = [
-    # '/Game/Carla/Maps/Town01', 
-    # '/Game/Carla/Maps/Town02', 
-    # '/Game/Carla/Maps/Town03', 
-    # '/Game/Carla/Maps/Town04', # has a lot of offroad actors
+    '/Game/Carla/Maps/Town01', 
+    '/Game/Carla/Maps/Town02', 
+    '/Game/Carla/Maps/Town03', 
+    '/Game/Carla/Maps/Town04', # has a lot of offroad actors
     '/Game/Carla/Maps/Town05', 
     '/Game/Carla/Maps/Town07', 
     # '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors
@@ -98,7 +98,7 @@ CAUTIOUS_IN_WEATHER_PROBABILITY = 0.6  # 60% drive more cautiously in bad weathe
 n_steps = 300
 timeout = 150.0
 for map in maps:
-    for ijk in range(5):
+    for ijk in range(3):
         client = carla.Client("localhost", 2000)
         client.set_timeout(timeout)
         client.load_world(map)

### 2025-10-06 graph_creator/graph_embeddings.py

@@ -55,13 +55,13 @@ def get_feature_dimensions(actor_type_mapping=None, edge_type_mapping=None):
         actor_type_mapping = ACTOR_TYPE_MAPPING
     if edge_type_mapping is None:
         edge_type_mapping = EDGE_TYPE_MAPPING
-    
-    # Node features: 1 (lon_speed) + num_actor_types (one-hot)
-    node_features = 1 + len(actor_type_mapping)
-    
+
+    # Node features: 1 (lon_speed) + num_actor_types (one-hot) + 2 (boolean attributes)
+    node_features = 1 + len(actor_type_mapping) + 2
+
     # Edge features: 1 (path_length) + num_edge_types (one-hot)
     edge_features = 1 + len(edge_type_mapping)
-    
+
     return node_features, edge_features
 
 def networkx_to_pyg(nx_graph, actor_type_mapping=None, edge_type_mapping=None):
@@ -92,27 +92,31 @@ def networkx_to_pyg(nx_graph, actor_type_mapping=None, edge_type_mapping=None):
     node_features = []
     for node in nodes:
         node_data = nx_graph.nodes[node]
-        
+
         # Extract lon_speed (continuous feature)
         lon_speed = node_data.get('lon_speed', 0.0)
-        
+
         # Extract actor_type and convert to one-hot
         actor_type = node_data.get('actor_type')
         if hasattr(actor_type, 'value'):  # Handle enum
             actor_type_str = actor_type.value
         else:
             actor_type_str = str(actor_type)
-        
+
         # Get actor type index
         actor_type_idx = actor_type_mapping.get(actor_type_str, 0)  # Default to 0 if unknown
-        
+
         # Create one-hot encoding for actor type
         num_actor_types = len(actor_type_mapping)
         actor_onehot = [0.0] * num_actor_types
         actor_onehot[actor_type_idx] = 1.0
-        
-        # Combine features: [lon_speed, actor_type_onehot...]
-        node_feature = [lon_speed] + actor_onehot
+
+        # Extract boolean attributes
+        lane_change = float(node_data.get('lane_change', False))  # Convert boolean to float
+        is_on_intersection = float(node_data.get('is_on_intersection', False))  # Convert boolean to float
+
+        # Combine features: [lon_speed, actor_type_onehot..., lane_change, is_on_intersection]
+        node_feature = [lon_speed] + actor_onehot + [lane_change, is_on_intersection]
         node_features.append(node_feature)
     
     # Convert to tensor

### 2025-10-08 carla/create_data_script_synchronous.py
Carla example data for Marius

@@ -97,287 +97,288 @@ CAUTIOUS_IN_WEATHER_PROBABILITY = 0.6  # 60% drive more cautiously in bad weathe
 
 n_steps = 300
 timeout = 150.0
-for map in maps:
-    for ijk in range(3):
-        client = carla.Client("localhost", 2000)
-        client.set_timeout(timeout)
-        client.load_world(map)
-        world = client.get_world()
-        clean_carla(world)
-        _ = world.tick()
-        # create the lane map graph and store it to file:
-        world_map = world.get_map()
-        map_g = MapGraph()
-        map_g = map_g.create_from_carla_map(world_map)
-        dt_specifier = str(datetime.now())
-        map_name = map.split("/")[-1]
-        scene_name = f"{dt_specifier}_{map_name}"
-        map_g.store_graph_to_file(f"carla/data/scene_{scene_name}_map_graph.pickle")
-        xodr = world_map.to_opendrive()
-        with open(f"carla/data/scene_{scene_name}_map.xodr", "w") as f:
-            f.write(xodr)
-        # Get the world and set synchronous mode
-        settings = world.get_settings()
-        settings.synchronous_mode = True  # Enable sync mode
-        settings.fixed_delta_seconds = 0.05  # Set fixed time step
-        world.apply_settings(settings)
-        tm = client.get_trafficmanager(8000)  # Port 8000
-        tm.set_synchronous_mode(True)  # Make TM sync with simulation
-        # Configure global traffic manager behavior for variation
-        # if SPEED_VARIATION_ENABLED: # throws error
-        #     # tm.set_global_percentage_speed_difference(-20.0)  # Base speed reduction
-        #     tm.vehicle_percentage_speed_difference(-20.0)  # Base speed reduction
-        if DISTANCE_VARIATION_ENABLED:
-            tm.set_global_distance_to_leading_vehicle(2.0)  # Base following distance
-        # Get blueprint library
-        blueprint_library = world.get_blueprint_library()
-        # Spawn vehicles
-        spawn_points = world.get_map().get_spawn_points()
-        random.shuffle(spawn_points)
-        # Track vehicle behavior states
-        vehicle_behaviors = {}
-        vehicle_slowdown_timers = {}
-        n_vehicles = random.randint(10, 60)
-        for i in range(n_vehicles):
-            spawn_point = spawn_points[i]
-            # Select vehicle type based on probabilities
-            if VEHICLE_TYPE_VARIATION_ENABLED:
[…]

### 2025-10-08 article/article.tex

@@
 
-The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.
-The model consists of multiple GINE convolutional layers, each containing:
+The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
+
+The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of randomly samples scenarios, the scenario with 
+closest embedding vector (euclidean distance as well as cosine distance) has been visualized. This is shown in Figure \ref{fig:plausibility_checks}. This 
+includes comparison between CARLA and Argoverse scenarios.
+
+As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
+different flavors, like for example:
+
 \begin{itemize}
-    \item Sequential multi-layer perceptrons (MLPs) with batch normalization and ReLU activations
-    \item Edge-aware message passing incorporating both node and edge features
-    \item Dropout regularization to prevent overfitting
+    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
+    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
+    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
+    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
+    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
 \end{itemize}
 
-The model employs a combination of three graph pooling strategies - mean, max, and sum pooling - to create a comprehensive graph-level representation.
+As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
+many other tasks, like for example clustering, anomaly detection, similarity search, etc.
+
+The results for this approach are shown in Section xxx.
+
+% The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.
+The model consists of multiple GINE convolutional layers, each containing:
+% \begin{itemize}
+%     \item Sequential multi-layer perceptrons (MLPs) with batch normalization and ReLU activations
+%     \item Edge-aware message passing incorporating both node and edge features
+%     \item Dropout regularization to prevent overfitting
+% \end{itemize}
+
+% The model employs a combination of three graph pooling strategies - mean, max, and sum pooling - to create a comprehensive graph-level representation.
 A projection head enables contrastive learning through InfoNCE loss, while an optional classification head supports supervised learning tasks.
 
-\subsection{Data Processing and Augmentation}
+% \subsection{Data Processing and Augmentation}
 
-The \texttt{GraphDataset} class handles the conversion from NetworkX graphs to PyTorch Geometric format through the \texttt{networkx\_to\_pyg} function. This conversion includes:
-\begin{itemize}
-    \item One-hot encoding of categorical node features (actor types: VEHICLE, PEDESTRIAN, CYCLIST, MOTORCYCLE)
-    \item One-hot encoding of edge types (neighbor\_vehicle, opposite\_vehicle, same\_lane, adjacent\_lane, following, intersection)
-    \item Preservation of continuous features such as longitudinal speed and path length
-\end{itemize}
+% The \texttt{GraphDataset} class handles the conversion from NetworkX graphs to PyTorch Geometric format through the \texttt{networkx\_to\_pyg} function. This conversion includes:
+% \begin{itemize}
+%     \item One-hot encoding of categorical node features (actor types: VEHICLE, PEDESTRIAN, CYCLIST, MOTORCYCLE)
+%     \item One-hot encoding of edge types (neighbor\_vehicle, opposite\_vehicle, same\_lane, adjacent\_lane, following, intersection)
+%     \item Preservation of continuous features such as longitudinal speed and path length
+% \end{itemize}
 
-Data augmentation is implemented through the \texttt{augment\_graph} function, which adds Gaussian noise to continuous features to create augmented versions for contrastive learning.
+% Data augmentation is implemented through the \texttt{augment\_graph} function, which adds Gaussian noise to continuous features to create augmented versions for contrastive learning.
 
-\subsection{Training Methodology}
+% \subsection{Training Methodology}
 
-The training process utilizes self-supervised contrastive learning with the following characteristics:
-\begin{itemize}
-    \item InfoNCE contrastive loss implemented in the \texttt{contrastive\_loss} function
-    \item Adaptive learning rate schedule with exponential decay (starting at 0.02, decaying by factor 0.75)
-    \item Multi-epoch training with 15 epochs per learning rate stage
-    \item Data from both CARLA simulator (4,442 graphs) and Argoverse 2.0 dataset (3,588 graphs)
-\end{itemize}
+% The training process utilizes self-supervised contrastive learning with the following characteristics:
+% \begin{itemize}
+%     \item InfoNCE contrastive loss implemented in the \texttt{contrastive\_loss} function
+%     \item Adaptive learning rate schedule with exponential decay (starting at 0.02, decaying by factor 0.75)
+%     \item Multi-epoch training with 15 epochs per learning rate stage
+%     \item Data from both CARLA simulator (4,442 graphs) and Argoverse 2.0 dataset (3,588 graphs)
+% \end{itemize}
 
-\subsection{Embedding Analysis and Visualization}
+% \subsection{Embedding Analysis and Visualization}
 
-The trained model generates 256-dimensional embeddings for each traffic scene graph. Analysis includes:
-\begin{itemize}
-    \item Dimensionality reduction using Principal Component Analysis (PCA) and t-distributed Stochastic Neighbor Embedding (t-SNE)
-    \item Visualization of embedding space distinguishing between CARLA and Argoverse data sources
-    \item Similarity-based scene retrieval using squared Euclidean distance in embedding space
-    \item Integration with visualization functions (\texttt{plot\_lane\_map\_advanced}, \texttt{add\_actors\_to\_map}, \texttt{add\_actor\_edges\_to\_map}) to display similar traffic scenes
-\end{itemize}
+% The trained model generates 256-dimensional embeddings for each traffic scene graph. Analysis includes:
+% \begin{itemize}
+%     \item Dimensionality reduction using Principal Component Analysis (PCA) and t-distributed Stochastic Neighbor Embedding (t-SNE)
+%     \item Visualization of embedding space distinguishing between CARLA and Argoverse data sources
+%     \item Similarity-based scene retrieval using squared Euclidean distance in embedding space
+%     \item Integration with visualization functions (\texttt{plot\_lane\_map\_advanced}, \texttt{add\_actors\_to\_map}, \texttt{add\_actor\_edges\_to\_map}) to display similar traffic scenes
+% \end{itemize}
 
-The implementation demonstrates successful learning of traffic scene representations, with the embedding space capturing meaningful similarities between scenes from different data sources while maintaining distinction between simulation and real-world data.
+% The implementation demonstrates successful learning of traffic scene representations, with the embedding space capturing meaningful similarities between scenes from different data sources while maintaining distinction between simulation and real-world data.
 

### 2025-10-11 carla/create_data_script_synchronous.py
get the argoverse thing running

@@ -138,7 +138,7 @@ for rep in range(3):
             # Track vehicle behavior states
             vehicle_behaviors = {}
             vehicle_slowdown_timers = {}
-            n_vehicles = random.randint(10, 60)
+            n_vehicles = random.randint(25, 65)
             for i in range(n_vehicles):
                 spawn_point = spawn_points[i]
                 # Select vehicle type based on probabilities

### 2025-10-12 article/article.tex
paper writing

@@
 
-\section{Analysing a traffic scene with agraph}
+\section{Analysing a traffic scene with a graph}
 
-\begin{itemize}
-    \item Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-    \item Two methodologies are proposed for this purpose:
-    \item One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-    \item The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-\end{itemize}
+Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
+Two methodologies are proposed for this purpose:
+One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
+The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
 
@@
 
-\begin{itemize}
-    \item There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-    \item Once an archetype is defined, a special property of graphs can be used. 
-    \item Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-    \item As the archetypes are not necessarily are involving a lot of actors, these are more like subsets of actual traffic scenes.
-    \item A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-    \item This situation can be represented by a graph with 3 nodes and 2 edges. 
-    \item In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-    \item This is an example of a subgraph isomorphism problem.
-    \item While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-    \item One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-    \item The strategy we are then applying is the following:
-    \item \begin{enumerate}
-            \item Define a set of subgraphs $S$ that are considered to be archetypes.
+There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
+Once an archetype is defined, a special property of graphs can be used. 
+Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
+As the archetypes are not necessarily are involving a lot of actors, these are more like subsets of actual traffic scenes.
+A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
+This situation can be represented by a graph with 3 nodes and 2 edges. 
+In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
+This is an example of a subgraph isomorphism problem.
+While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
+One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
+The strategy we are then applying is the following:
+
+\begin{enumerate}
+            \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
             \item Define which node and edge attributes are considered for the isomorphism check.
@@
 \end{enumerate}
-    \item This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-    \item Then going upwards to different datasets, it is checked, if the archetype is present.
-    \item Also, follow up analysis of the created coverage dataframe can be performed. For example,
-    \item \begin{itemize}
+
+This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
+Then going upwards to different datasets, it is checked, if the archetype is present.
+Also, follow up analysis of the created coverage dataframe can be performed. For example,
+
+\begin{itemize}
     \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
@@
     \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \end{itemize}
 \end{itemize}
@@
 many other tasks, like for example clustering, anomaly detection, similarity search, etc.
+
+And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
+Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
+if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
 

### 2025-10-12 carla/create_data_script_synchronous.py

@@ -39,7 +39,7 @@ maps = [
     '/Game/Carla/Maps/Town01', 
     '/Game/Carla/Maps/Town02', 
     '/Game/Carla/Maps/Town03', 
-    '/Game/Carla/Maps/Town04', # has a lot of offroad actors
+    '/Game/Carla/Maps/Town04',
     '/Game/Carla/Maps/Town05', 
     '/Game/Carla/Maps/Town07', 
     # '/Game/Carla/Maps/Town04_Opt', # has a lot of offroad actors

### 2025-10-12 carla/create_data_script_synchronous.py
paper writing and citation clean up

@@ -99,7 +99,7 @@ n_steps = 300
 timeout = 150.0
 for rep in range(3):
     for map in maps:
-        for ijk in range(2):
+        for ijk in range(1):
             client = carla.Client("localhost", 2000)
             client.set_timeout(timeout)
             client.load_world(map)
@@ -379,6 +379,6 @@ for rep in range(3):
             tracks_df["map"] = world.get_map().name
             tracks_df["scene_id"] = dt_specifier
             tracks_df.to_parquet(f"carla/data/scene_{scene_name}_tracks.parquet")
-            print(f"scene {scene_name} saved")
+            print(f"scene {scene_name} saved, rep {rep}")
             time.sleep(5)
 

### 2025-10-12 article/article.tex

@@
 
+test2
+
 \section{Introduction}
@@
     \item In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-    \item There are already other graph-based approaches for analysing and representing traffic scenes, see for example \cite{behr2021graph}.
+    \item There are already other graph-based approaches for analysing and representing traffic scenes.
     \item However, the work in that paper is not specifically focused on coverage analysis.
@@
 \end{itemize}
-
-\cite{wagner2022odd}
-
-\cite{arzamasov2021data}
 
@@
 
-\cite{berger2020survey}
-
 \cite{iso21448}
@@
 
-\cite{behr2021graph}
+still to be added: https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf
+
+still to be added: https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf
 
@@
 
-\cite{bagheri2020ontology}
-
-\cite{riedmaier2020realistic}
+\cite{bagschik2018ontologybasedscenecreation}
 
 \cite{fremont2020scenic}
-
-\cite{pek2019generating}
 

### 2025-10-13 article/article.tex

--- 2025-10-12 article/article.tex
@@
 
+test3
+
 \section{Introduction}
 
-\begin{itemize}
-    \item In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-    \item In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-    \item See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-    \item In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-    \item There are already other graph-based approaches for analysing and representing traffic scenes, see for example \cite{behr2021graph}.
-    \item However, the work in that paper is not specifically focused on coverage analysis.
-    \item Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-    \item This paper is structured as follows:
-    \begin{itemize}
-        \item In the first section, xxx
-    \end{itemize}
-\end{itemize}
+In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
+In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
+See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
+In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
+There are already other graph-based approaches for analysing and representing traffic scenes.
+However, the work in that paper is not specifically focused on coverage analysis.
+Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
 
-\cite{wagner2022odd}
+This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
+The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
+using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
+chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
+Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
+followed by a short summary and outlook.
 
-\cite{arzamasov2021data}
-
-\section{exsting coverage and analysis approaches}
+\section{Existing coverage and analysis approaches}
+\label{chapter:existing_approaches}
 
@@
 \section{Defining a traffic scene graph}
+\label{chapter:defining_a_traffic_scene_graph}
 
-\cite{behr2021graph}
+still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
+
+still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
 
@@
 
-\cite{bagheri2020ontology}
-
-\cite{riedmaier2020realistic}
+\cite{bagschik2018ontologybasedscenecreation}
 
@@
 
-\cite{pek2019generating}
+\subsection{time based graph representations}
+The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
+extension is to create a graph which contains information for multiple timesteps.
 
-\subsection{time based graph representations}
+\begin{itemize}
+    \item create a graph for multiple timesteps
+    \item create a joint set of nodes. 
+    \item For each time step, determine the corresponding edges.
+    \item Add all edges to the graph, resulting in a di-graph, which can contain for a pair of nodes as many edges as there are time steps.
+\end{itemize}
+
+This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
+changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
+to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
+if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
+non-equal dimensions of tensors for node attributes.
+
+The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
+the focus in this paper will be on single timestep graphs.
 
@@
 \section{Create subgraphs for coverage analysis}
+\label{chapter:create_subgraphs_for_coverage_analysis}
 
@@
 \section{Implementation of Graph Embeddings for Traffic Scene Analysis}
+\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
 
@@
 \section{Application}
+\label{chapter:application}
 

### 2025-10-17 article/article.tex
subgraph isomorphism plots

@@
 \usepackage{graphicx}
-
+\usepackage{booktabs}
 \title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}
@@
 
-test3
+test4
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
+    \caption{Coverage barchart for the manually defined coverage scenarios for CARLA.}
+    \label{fig:carla_coverage_barchart_manual_scenarios}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
+    \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
+    \label{fig:carla_coverage_barchart_manual_scenarios}
+\end{figure}
+
+\begin{table}[htbp]
+    \caption{My Table}
+    \label{tab:mytable}
+    \begin{tabular}{rrrrr}
+    \toprule
+    scn 1 & scn 2 & scn 3 & scn 4 & size \\
+    \midrule
+    False & False & False & False & 613 \\
+    False & False & True & False & 3 \\
+    True & True & False & True & 377 \\
+    True & True & True & True & 7 \\
+    \bottomrule
+    \end{tabular}
+\end{table}
 % \cite{goodfellow2016deep} % needed?

### 2025-10-17 graph_creator/SubgraphIsomorphismChecker.py
subgraph_isom_histograms

@@ -46,5 +46,6 @@ class IsomorphicGrapCoverageCounter:
             self.cov_data["degree"].append(sum(dict(ag_nx.degree()).values()) / len(ag_nx.nodes()))
             self.cov_data["density"].append(nx.density(ag_nx))
             self.cov_data["diameter"].append(nx.diameter(ag_nx))
+            self.cov_data["path"] = graph_path
 
         self.cov_data_df = pd.DataFrame(self.cov_data)

### 2025-10-17 article/article.tex
subgraph_isom_histograms
small updates

@@
     \caption{Coverage barchart for the manually defined coverage scenarios for CARLA.}
-    \label{fig:carla_coverage_barchart_manual_scenarios}
+    \label{fig:subgraph_isomorphism_carla_coverage_barchart}
 \end{figure}
@@
     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-    \label{fig:carla_coverage_barchart_manual_scenarios}
+    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
 \end{figure}
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.95\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
+    \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
+    \label{fig:combined_distributions_plots_speed_distance_carla}
+\end{figure}
+
 \section{Summary}

### 2025-10-17 graph_creator/graph_embeddings.py
add architecture diagram

@@ -32,7 +32,6 @@ ACTOR_TYPE_MAPPING = {
     'PEDESTRIAN': 1,
     'CYCLIST': 2,
     'MOTORCYCLE': 3,
-    # Add more types as needed
 }
 
 EDGE_TYPE_MAPPING = {
@@ -42,7 +41,6 @@ EDGE_TYPE_MAPPING = {
     'adjacent_lane': 3,
     'following': 4,
     'intersection': 5,
-    # Add more edge types as needed
 }
 
 

### 2025-10-17 article/article.tex

@@
 \usepackage{booktabs}
+\usepackage{svg}
+
 \title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}
@@
 
-The exact architecture of the model is shown in Figure \ref{fig:model_architecture}. The features used are the actor type (as a one hot encoding),
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
+    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
+    \label{fig:graph_gine_architecture}
+\end{figure}
+
+The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features used are the actor type (as a one hot encoding),
 the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its lane since the last timestep (boolean) for the nodes.

### 2025-10-18 article/article.tex
last changes
create first few embedding plots

@@
 \title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}
-\author{xxx}
+\author{Thomas Mühlenstädt and Marius Bause}
 \date{\today}
@@
 
-\section{Implementation of Graph Embeddings for Traffic Scene Analysis}
+\section{Graph Embeddings for Traffic Scene Analysis}
 \label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
@@
 
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its lane since the last timestep (boolean) for the nodes.
+The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
+used are the actor type (as a one hot encoding),
+the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
+lane since the last timestep (boolean) for the nodes.
 For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
@@
 
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance as well as cosine distance) has been visualized. This is shown in Figure \ref{fig:plausibility_checks}. This 
+Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
+For each batch, two correlated views of every graph were created by 
+perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
+longitudinal speed and edge path length. A four-layer GINE encoder 
+(hidden width 96) produced graph-level representations via the concatenation of 
+mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
+The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
+Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
+This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
+and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
+\caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
+\label{fig:train_test_graph_embeddings_loss_plot}
+\end{figure}
+
+The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
+randomly samples scenarios, the scenario with 
+closest embedding vector (euclidean distance as well as cosine distance) has been visualized. 
+This is shown in Figure \ref{fig:plausibility_checks}. This 
 includes comparison between CARLA and Argoverse scenarios.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
+    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
+    \label{fig:embedding_space}
+\end{figure}
 

### 2025-10-19 article/article.tex

@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
+
+    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
+
+    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
+
+    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
+
+    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
+    \caption{Plausibility checks showing graph embedding comparisons.
+For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
+    \label{fig:plausibility_checks}
+\end{figure}
+
 \end{document}

### 2025-10-20 article/article.tex
start writing the reference section

@@
 
-\cite{pegasus2019method}
+There exist a larger 
+The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
+validation of highly automated driving functions, addressing the impracticality of traditional 
+distance-based testing approaches. The core of the method is a six-layer model used to 
+systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
+This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
+continuous test space into discrete, manageable "logical scenarios". By systematically 
+parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
+the methodology aims to ensure the completeness of relevant test runs and provides a structured 
+foundation for arguing that the automated system has been adequately tested across its entire operational design 
+domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
+for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
 
-\cite{Ries2021traj_clustering}
+The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
+in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
+They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
+clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
+scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
+supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
+not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
 
-\cite{ORAD2021taxonomy}
+In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
+situation, and scenario—are inconsistently defined in the literature, which complicates the 
+development, testing and validation of driving-automation modules. They propose clear 
+definitions: a scene is a snapshot of the environment including dynamic and static 
+elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
+behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
+values; and a scenario is a temporal development of several scenes in sequence, involving 
+actions/events and goals/values. The paper also provides example implementations of these 
+definitions in the context of automated-vehicle systems and how they interface with perception, 
+planning and control, and testing.
 
-\cite{Ulbrich2015scene}
+Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
+provides many foundational terms in the context of autonomous driving.
+It provides a functional taxonomy and clear definitions for key terms related to 
+driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
+levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
+the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
+important related terms such as the Dynamic Driving Task (DDT), Operational Design 
+Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
+among others. The 2021-04 revision (superseding the 2018 version) was developed in 
+collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
+(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
+it does not prescribe specifications or impose performance requirements for DAS.
 
-\cite{DBLP:journals/corr/abs-1801-08598}
+The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
+validating automated-driving systems across different development phases. It 
+introduces three abstraction levels of scenarios—functional, logical, and 
+concrete—and discusses how these can be transformed for use in testing. The authors note that 
+existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
+systematic way to determine meaningful test coverage, highlighting coverage analysis as 
+an open challenge in scenario-based validation.
 
-\cite{Ammann_Offutt_2008}
+In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on autonomous driving, the authors provide an 
+introduction to software testing, which is a 
+good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
+different types of coverage, such as statement coverage, branch coverage, condition coverage, 
+and decision coverage. They also discuss the different techniques for measuring coverage, such as 
+branch coverage, condition coverage, and decision coverage.
 

### 2025-10-21 article/article.tex
paper writing

@@
 
-There exist a larger 
+There exist a large amount of literature on coverage analysis in the context of autonomous driving.
+In the following, some of the most relevant approaches are discussed.
+
 The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
@@
 
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on autonomous driving, the authors provide an 
-introduction to software testing, which is a 
+In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
+autonomous driving, the authors provide an introduction to software testing, which is a 
 good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
@@
 
-\subsection{time based graph representations}
+\subsection{Time based graph representations}
 The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps.
-
-\begin{itemize}
-    \item create a graph for multiple timesteps
-    \item create a joint set of nodes. 
-    \item For each time step, determine the corresponding edges.
-    \item Add all edges to the graph, resulting in a di-graph, which can contain for a pair of nodes as many edges as there are time steps.
-\end{itemize}
+extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
+first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
+Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
+While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
+there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
+time stamp of the edge.
 
@@
 Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily are involving a lot of actors, these are more like subsets of actual traffic scenes.
+As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
 A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
@@
 \end{itemize}
+
+In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
+from a dataset of real world traffic observation data.
+
+xxxxxx
+
+TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
+
+xxxxxx
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-\caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-\label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance as well as cosine distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-The results for this approach are shown in Section xxx.
-
 % The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.
@@
 
+TODO MARIUS: Describe the Argoverse 2.0 dataset.
+
 \subsection{Carla}
@@
 
+\subsection{Subgraph Isomorphism Coverage Analysis}
+
 \begin{figure}[h]
@@
 
+\subsection{Graph Embeddings Coverage Analysis}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
+    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
+    \label{fig:train_test_graph_embeddings_loss_plot}
+\end{figure}
+
+The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
+randomly samples scenarios, the scenario with 
+closest embedding vector (euclidean distance as well as cosine distance) has been visualized. 
+This is shown in Figure \ref{fig:plausibility_checks}. This 
+includes comparison between CARLA and Argoverse scenarios.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
+    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
+    \label{fig:embedding_space}
+\end{figure}
+
+As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
+different flavors, like for example:
+
+\begin{itemize}
+    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
+    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
+    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
+    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
+    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
+\end{itemize}
+
+As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
+many other tasks, like for example clustering, anomaly detection, similarity search, etc.
+
+And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
+Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
+if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
+
+The results for this approach are shown in Section xxx.
+
 \section{Summary}

### 2025-10-23 article/article.tex
update arch plot and finish ref section

@@
 
+In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
+in an object oriented manner, with the intention to have a clearly defined and implementable
+structures for scenarios, delivering a clear linkage to coverage arguments.
+
 The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
@@
 
-\cite{foretellix2019}
+A much more focussed example of coverage analysis in the context of autonomous driving is provided 
+by \cite{foretellix2019} in a blog 
+post, discussing topics like items, parameters and 
+coverage buckets and performance metrics like time to collision and error collections.
 
-\cite{deGelder2022ontology}
+The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
+production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
+as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
+The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
+superior safety through real-world driving, which would necessitate billions of test kilometers. 
+This validation gap is presented as a fundamental problem of coverage analysis, where the
+technical system must now demonstrably cover the vast operational domain previously managed by the 
+human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
+test case generation methodologies, such as critical scenario identification, and the extensive
+ use of validated simulation-based tools to achieve sufficient test coverage efficiently.
 
-\cite{wachenfeld2016release}
-
-\cite{iso21448}
-
-\cite{ul4600}
+Well known standards for coverage analysis in the context of autonomous driving are the 
+ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
+The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
+including automated driving systems.
+It defines a framework for coverage analysis, including the definition of coverage criteria and 
+the measurement of coverage.
+The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
+It defines a framework for coverage analysis, including the definition of coverage criteria and the 
+measurement of coverage.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
+    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
+    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
     \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}

### 2025-11-11 graph_creator/ActorGraph.py
add argoverse subgraph isomorphism plots

@@ -257,6 +257,13 @@ class ActorGraph:
             delta_timestep_s=delta_timestep_s,
         )
 
+        instance.actor_components = {}
+        # print("instance.actor_graphs.keys(): ", instance.actor_graphs.keys())
+        for key, value in instance.actor_graphs.items():
+            components = list(nx.weakly_connected_components(value))
+            subgraphs = [value.subgraph(c).copy() for c in components]
+            instance.actor_components[key] = subgraphs
+
         return instance
 
     @classmethod

### 2025-11-12 graph_creator/SubgraphIsomorphismChecker.py
debug subgIsom nb

@@ -28,10 +28,11 @@ class IsomorphicGrapCoverageCounter:
         self.cov_data["degree"] = []
         self.cov_data["density"] = []
         self.cov_data["diameter"] = []
+        self.cov_data["path"] = []
 
 
     def count_isomorphic_graphs(self):
-        for graph_path in tqdm(self.graph_paths):
+        for graph_path in tqdm(self.graph_paths, desc="Checking isomorphic graphs"):
             with open(graph_path, "rb") as file:
                 ag_nx = pickle.load(file)
 
@@ -46,6 +47,6 @@ class IsomorphicGrapCoverageCounter:
             self.cov_data["degree"].append(sum(dict(ag_nx.degree()).values()) / len(ag_nx.nodes()))
             self.cov_data["density"].append(nx.density(ag_nx))
             self.cov_data["diameter"].append(nx.diameter(ag_nx))
-            self.cov_data["path"] = graph_path
+            self.cov_data["path"].append(graph_path)
 
         self.cov_data_df = pd.DataFrame(self.cov_data)

### 2025-11-18 article/article.tex
some small changes to get it all clean.
subgraph iso app part writing start

@@
 \usepackage{svg}
+\usepackage{subcaption}
 
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving}
+\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
+
 \author{Thomas Mühlenstädt and Marius Bause}
@@
 
-test4
+tbd
+
+% $$\hat{r} = \frac{N}{E}$$
+
+% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
+% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
+
+% $$LL_{r} = LL_{\lambda} / E$$
+% $$UL_{r} = UL_{\lambda} / E$$
 
@@
 
-% The core component is the \texttt{TrainableGraphGINE} class, which implements a Graph Isomorphism Network with Edge features (GINE) architecture for generating graph-level embeddings.
-The model consists of multiple GINE convolutional layers, each containing:
-% \begin{itemize}
-%     \item Sequential multi-layer perceptrons (MLPs) with batch normalization and ReLU activations
-%     \item Edge-aware message passing incorporating both node and edge features
-%     \item Dropout regularization to prevent overfitting
-% \end{itemize}
-
-% The model employs a combination of three graph pooling strategies - mean, max, and sum pooling - to create a comprehensive graph-level representation.
-A projection head enables contrastive learning through InfoNCE loss, while an optional classification head supports supervised learning tasks.
-
-% \subsection{Data Processing and Augmentation}
-
-% The \texttt{GraphDataset} class handles the conversion from NetworkX graphs to PyTorch Geometric format through the \texttt{networkx\_to\_pyg} function. This conversion includes:
-% \begin{itemize}
-%     \item One-hot encoding of categorical node features (actor types: VEHICLE, PEDESTRIAN, CYCLIST, MOTORCYCLE)
-%     \item One-hot encoding of edge types (neighbor\_vehicle, opposite\_vehicle, same\_lane, adjacent\_lane, following, intersection)
-%     \item Preservation of continuous features such as longitudinal speed and path length
-% \end{itemize}
-
-% Data augmentation is implemented through the \texttt{augment\_graph} function, which adds Gaussian noise to continuous features to create augmented versions for contrastive learning.
-
-% \subsection{Training Methodology}
-
-% The training process utilizes self-supervised contrastive learning with the following characteristics:
-% \begin{itemize}
-%     \item InfoNCE contrastive loss implemented in the \texttt{contrastive\_loss} function
-%     \item Adaptive learning rate schedule with exponential decay (starting at 0.02, decaying by factor 0.75)
-%     \item Multi-epoch training with 15 epochs per learning rate stage
-%     \item Data from both CARLA simulator (4,442 graphs) and Argoverse 2.0 dataset (3,588 graphs)
-% \end{itemize}
-
-% \subsection{Embedding Analysis and Visualization}
-
-% The trained model generates 256-dimensional embeddings for each traffic scene graph. Analysis includes:
-% \begin{itemize}
-%     \item Dimensionality reduction using Principal Component Analysis (PCA) and t-distributed Stochastic Neighbor Embedding (t-SNE)
-%     \item Visualization of embedding space distinguishing between CARLA and Argoverse data sources
-%     \item Similarity-based scene retrieval using squared Euclidean distance in embedding space
-%     \item Integration with visualization functions (\texttt{plot\_lane\_map\_advanced}, \texttt{add\_actors\_to\_map}, \texttt{add\_actor\_edges\_to\_map}) to display similar traffic scenes
-% \end{itemize}
-
-% The implementation demonstrates successful learning of traffic scene representations, with the embedding space capturing meaningful similarities between scenes from different data sources while maintaining distinction between simulation and real-world data.
-
 \section{Application}
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-    \caption{Coverage barchart for the manually defined coverage scenarios for CARLA.}
-    \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-\end{figure}
+\begin{itemize}
+    \item We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
+    \item The archetypes used here are defined manually and described in table xxx.
+The chosen archetypes are typical traffic situations like e.g.
+lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
+    \item Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
+    \item The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}.
+    \item In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform.
+    \item Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
+    \item Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
+\end{itemize}
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
+    \begin{subfigure}[b]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
+        \caption{CARLA coverage scenarios}
+        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
+        \caption{Argoverse coverage scenarios}
+        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
+    \end{subfigure}
+    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
+    \label{fig:subgraph_isomorphism_coverage_barcharts}
+\end{figure}
+
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
+%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
+%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
+% \end{figure}
+
+\begin{figure}[h]
+    \centering
+    \begin{subfigure}[b]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
+        \caption{Agreement matrix for CARLA.}  
         \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
+        \caption{Agreement matrix for Argoverse.}  
+        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
+    \end{subfigure}
+    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
+    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.95\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
+    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
+    \caption{Parameter distribution for speed and path length for CARLA.}
     \label{fig:combined_distributions_plots_speed_distance_carla}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
+    \caption{Parameter distribution for speed and path length for Argoverse.}
+    \label{fig:combined_distributions_plots_speed_distance_argo}
 \end{figure}

### 2025-11-19 article/article.tex
start summary writing

@@
 
-\begin{itemize}
-    \item We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-    \item The archetypes used here are defined manually and described in table xxx.
+We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
+The archetypes used here are defined manually and described in table xxx.
 The chosen archetypes are typical traffic situations like e.g.
 lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-    \item Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-    \item The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}.
-    \item In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform.
-    \item Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-    \item Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-\end{itemize}
+Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
+The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
+In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
+Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
+Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
+A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
+Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
+The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
+
+TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
+
+A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
+Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
+The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
+Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
 
@@
 
-\begin{table}[htbp]
-    \caption{My Table}
-    \label{tab:mytable}
-    \begin{tabular}{rrrrr}
-    \toprule
-    scn 1 & scn 2 & scn 3 & scn 4 & size \\
-    \midrule
-    False & False & False & False & 613 \\
-    False & False & True & False & 3 \\
-    True & True & False & True & 377 \\
-    True & True & True & True & 7 \\
-    \bottomrule
-    \end{tabular}
-\end{table}
-% \cite{goodfellow2016deep} % needed?
+% \begin{table}[htbp]
+%     \caption{My Table}
+%     \label{tab:mytable}
+%     \begin{tabular}{rrrrr}
+%     \toprule
+%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
+%     \midrule
+%     False & False & False & False & 613 \\
+%     False & False & True & False & 3 \\
+%     True & True & False & True & 377 \\
+%     True & True & True & True & 7 \\
+%     \bottomrule
+%     \end{tabular}
+% \end{table}
+% % \cite{goodfellow2016deep} % needed?
 
@@
 randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance as well as cosine distance) has been visualized. 
+closest embedding vector (euclidean distance) has been visualized. 
 This is shown in Figure \ref{fig:plausibility_checks}. This 
@@
 
-The results for this approach are shown in Section xxx.
+\section{Summary}
 
-\section{Summary}
+\begin{itemize}
+    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
+    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
+    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
+    \item Then it is checked, which archetypes are present in a traffic scene and how often.
+    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
+    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
+    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
+    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
+Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
+    \item 
+    \item 
+\end{itemize}
 

### 2025-11-19 article/article.tex
summary writing

@@
     \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
+    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios archetypes and different driving situations.
+    \item Besides the definition of the actor graph, no specific handling of different types of traffic situations is necessary.
     \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
 Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
+    \item Also, the presented approach easily allows for different number of actors in the traffic scene without excluding actors or only focussing on a subset of actors.
 \end{itemize}

### 2025-11-19 article/article.tex
summary writing

@@
     \item Also, the presented approach easily allows for different number of actors in the traffic scene without excluding actors or only focussing on a subset of actors.
+    \item Next steps in the same line of this research will be to further optimize the actor graph to include several time steps into one graph. 
+    \item Also, automatically extracting archetypes from real world traffic observation data will be investigated.
+    \item The source code for this research is available at \url{https://github.com/tmuehlen80/graph_coverage}.
 \end{itemize}

### 2025-11-22 graph_creator/ActorGraph.py
fix issue with opposing actors

@@ -506,8 +506,17 @@ class ActorGraph:
                 else:  # Backward direction (negative path_length)
                     max_distance = max_distance_opposite_backward_m
                 
+                # Check path length constraint
                 if abs(path_length) < max_distance:
-                    return ("opposite_vehicle", path_length)
+                    # Additional validation: Check actual Euclidean distance
+                    # The lane-based path_length can be misleading when lanes are geometrically far apart
+                    pos_A = self.track_xyz_pos_dict[track_id_A][t]
+                    pos_B = self.track_xyz_pos_dict[track_id_B][t]
+                    euclidean_dist = np.sqrt((pos_A.x - pos_B.x)**2 + (pos_A.y - pos_B.y)**2)
+                    
+                    # Only create opposite relation if Euclidean distance is also within limit
+                    if euclidean_dist < max_distance:
+                        return ("opposite_vehicle", path_length)
 
         return None
 
@@ -1121,3 +1130,241 @@ class ActorGraph:
             plt.savefig(save_path)
         else:
             plt.show()
+
+
+def main():
+    """
+    Debug function to investigate the opposing actors distance issue at timestep 9
+    for the 20:53:41.635057_Town01 example.
+    """
+    import pandas as pd
+    from pathlib import Path
+    from graph_creator.MapGraph import MapGraph
+    
+    # Setup paths
+    work_dir = Path(__file__).parent.parent
+    scene_id = "2025-10-05 20:53:41.635057_Town01"
+    
+    print(f"Loading scene: {scene_id}")
+    
+    # Load data
+    tracks_path = work_dir / f"example_data/carla/scene_{scene_id}_tracks.parquet"
+    map_path = work_dir / f"example_data/carla/scene_{scene_id}_map_graph.pickle"
+    
+    print(f"Loading tracks from: {tracks_path}")
+    tracks = pd.read_parquet(tracks_path)
+    tracks['road_lane_id'] = tracks.road_id.astype(str) + '_' + tracks.lane_id.astype(str)
+    
+    print(f"Loading map from: {map_path}")
+    g_map = MapGraph()
+    g_map.read_graph_from_file(str(map_path))
+    
+    # Create ActorGraph with the parameters from the notebook
+    print("\nCreating ActorGraph...")
+    ag = ActorGraph()
[…]

### 2025-11-22 graph_creator/ActorGraph.py
bug euclidian - distance in discovery phase. Update vis with actor labels

@@ -381,27 +381,31 @@ class ActorGraph:
 
     def _has_path_within_distance(self, G, source, target, max_distance):
         """
-        Check if there's a path from source to target with at most max_distance nodes.
+        Check if there's a path from source to target with at most max_distance edges.
         
         Args:
             G: NetworkX graph
             source: Source node
             target: Target node
-            max_distance: Maximum number of nodes in the path
+            max_distance: Maximum number of edges in the path
             
         Returns:
-            True if there's a path with at most max_distance nodes, False otherwise
+            True if there's a path with at most max_distance edges, False otherwise
         """
         if source == target:
             return True
         
-        if max_distance <= 1:
+        if max_distance <= 0:
+            return False
+        
+        if max_distance == 1:
             return G.has_edge(source, target)
         
         # Use BFS to find shortest path
         try:
             path = nx.shortest_path(G, source, target)
-            return len(path) <= max_distance
+            # len(path) - 1 is the number of edges
+            return len(path) - 1 <= max_distance
         except nx.NetworkXNoPath:
             return False
 
@@ -411,6 +415,9 @@ class ActorGraph:
                                      max_distance_neighbor_forward_m, max_distance_opposite_forward_m, max_distance_opposite_backward_m):
         """
         Find the best relation from actor A to actor B.
+        
+        Discovery phase: finds all potential relations based on distance limits.
+        Does NOT check max_node_distance_* parameters - those are used in construction phase only.
         
         Args:
             track_id_A: ID of actor A
@@ -456,7 +463,15 @@ class ActorGraph:
                     - self.track_s_value_dict[track_id_A][t]
                 )
                 if path_length < max_distance_lead_veh_m:
-                    return ("following_lead", path_length)
+                    # Additional validation: Check actual Euclidean distance
+                    # Lane-based path_length can be misleading on curved roads
+                    pos_A = self.track_xyz_pos_dict[track_id_A][t]
[…]

### 2025-11-22 graph_creator/plot_graphs.py

@@ -1,302 +1,400 @@
-
 import matplotlib.pyplot as plt
 import matplotlib.patches as patches
-from shapely.geometry import Polygon
+from shapely.geometry import Polygon, Point as ShapelyPoint
 import numpy as np
 
-def plot_lane_map(graph, figsize=(15, 10), intersection_color='red', lane_color='blue', alpha=0.6):
-   """
-   Plot a map showing lanes as polygons from a NetworkX graph.
-   
-   Parameters:
-   - graph: NetworkX graph with node_info containing lane_polygon
-   - figsize: Figure size tuple
-   - intersection_color: Color for intersection lanes
-   - lane_color: Color for regular lanes
-   - alpha: Transparency level
-   """
-   fig, ax = plt.subplots(figsize=figsize)
-   
-   # Lists to store all coordinates for setting axis limits
-   all_x_coords = []
-   all_y_coords = []
-   
-   # Plot each lane polygon
-   for node_id, node_data in graph.nodes(data=True):
-       node_info = node_data['node_info']
-       lane_polygon = node_info.lane_polygon
-       lane_id = node_info.lane_id
-       is_intersection = node_info.is_intersection
-       
-       # Extract coordinates from shapely polygon
-       if lane_polygon and hasattr(lane_polygon, 'exterior'):
-           coords = list(lane_polygon.exterior.coords)
-           x_coords = [coord[0] for coord in coords]
-           y_coords = [coord[1] for coord in coords]
-           
-           # Add to all coordinates for axis limits
-           all_x_coords.extend(x_coords)
-           all_y_coords.extend(y_coords)
-           
-           # Choose color based on intersection status
-           color = intersection_color if is_intersection else lane_color
-           
-           # Create polygon patch
-           polygon_patch = patches.Polygon(coords, 
-                                         facecolor=color, 
-                                         edgecolor='black', 
-                                         alpha=alpha,
-                                         linewidth=0.5)
-           ax.add_patch(polygon_patch)
-           
-           # Add lane ID as text at centroid
-           centroid = lane_polygon.centroid
-           ax.text(centroid.x, centroid.y, lane_id, 
-                  fontsize=8, ha='center', va='center',
-                  bbox=dict(boxstyle='round,pad=0.3', facecolor='white', alpha=0.8))
-   
-   # Set axis limits and properties
-   if all_x_coords and all_y_coords:
[…]

### 2025-11-23 article/sections/abstract.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
 \section{Abstract}
@@
 % $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-23 article/sections/application.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
 \section{Application}
@@
 
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
 \begin{figure}[h]
@@
 \end{figure}
-
-\end{document}

### 2025-11-23 article/sections/create_subgraphs.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
 \section{Create subgraphs for coverage analysis}
@@
 xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-23 article/sections/defining_traffic_scene_graph.tex
article tex seperated. Graph construction part added

\section{Defining a traffic scene graph}
\label{chapter:defining_a_traffic_scene_graph}

still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}

still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}

\cite{newman2010networks}

\cite{bagschik2018ontologybasedscenecreation}

\cite{fremont2020scenic}

\subsection{Map Graph Construction}
\label{subsec:map_graph}

The map graph serves as the foundational structure for traffic scene graph construction, encoding the 
spatial relationships between lanes in the road network. It is a directed multigraph 
$G_{\text{map}} = (V_{\text{map}}, E_{\text{map}})$ where nodes represent individual lanes and edges 
encode three fundamental spatial relationships between lanes.

Each node $v \in V_{\text{map}}$ represents a lane segment and stores geometric and semantic 
information including:
\begin{itemize}
    \item Lane boundaries (left and right boundaries as polylines)
    \item Lane length
    \item Lane centerline
    \item Intersection status (whether the lane is within an intersection)
    \item Road type and lane type information
\end{itemize}

The map graph defines three edge types that capture different spatial relationships:

\begin{enumerate}
    \item \textbf{Following edges} ($e_{\text{following}}$): Connect lanes that form a continuous 
          path in the same direction.

    \item \textbf{Neighbor edges} ($e_{\text{neighbor}}$): Connect adjacent lanes traveling in 
          the same direction.

    \item \textbf{Opposite edges} ($e_{\text{opposite}}$): Connect lanes traveling in opposite 
          directions. 
\end{enumerate}

The map graph construction algorithm processes the original map data from Argoverse or CARLA  
to extract these relationships. For intersection detection, lanes are analyzed for geometric 
overlap, and all overlapping lanes are marked as intersection lanes to ensure proper handling of 
complex road geometries.

%TODO: add figure

\subsection{Actor Graph Construction}
\label{subsec:actor_graph}

The actor graph $G_{\text{actor}} = (V_{\text{actor}}, E_{\text{actor}})$ represents the dynamic 
relationships between vehicles and other actors in a traffic scene at a specific timestep. This 
graph is constructed using a two-phase algorithm that separates relation discovery from graph 
construction. The second step serves to reduce the number of relations in the graph for computational tractability, 
and enables the representation of relationships between actors through indirect paths involving multiple intermediate nodes.

Each node $v \in V_{\text{actor}}$ represents an actor (vehicle, pedestrian, etc.) and stores the 
following attributes:

\begin{itemize}
    \item \textbf{Primary lane ID}: The lane on which the actor is primarily located
    \item \textbf{Lane IDs}: List of all lanes the actor occupies (for actors spanning multiple 
          lanes)
    \item \textbf{Longitudinal position} ($s$): Position along the lane's centerline
    \item \textbf{3D position} ($x, y, z$): Cartesian coordinates
    \item \textbf{Longitudinal speed}: Speed along the lane direction
    \item \textbf{Actor type}: Classification (vehicle, pedestrian, etc.)
    \item \textbf{Lane change indicator}: Boolean flag indicating if the actor changed lanes from 
          the previous timestep
\end{itemize}

Note that the node structure is extensible---users can add additional attributes such as 
acceleration, heading angle, or vehicle dimensions as needed for their coverage model.

Each edge $e \in E_{\text{actor}}$ represents a relationship between two actors and contains:

\begin{itemize}
    \item \textbf{Edge type}: One of four relation types between actors (leading, following, neighbor, opposite)
    \item \textbf{Path length}: Distance in meters along the lane network 
\end{itemize}

Similar to nodes, edge attributes can be extended with additional information such as relative 
velocities, time-to-collision, or interaction probabilities.

\label{subsubsec:relation_types}

The actor graph defines four types of relationships between actors, ordered by semantic hierarchy:

\begin{enumerate}
    \item \textbf{Following/Leading} (\texttt{following\_lead}, \texttt{leading\_vehicle}): 
          Represents longitudinal relationships where one actor follows another in the same 
          direction and lane.

    \item \textbf{Neighbor} (\texttt{neighbor\_vehicle}): Represents lateral relationships 
          between actors on lanes traveling in the same direction. The actors need not be on 
          immediately adjacent lanes.

    \item \textbf{Opposite} (\texttt{opposite\_vehicle}): Represents relationships between actors 
          on lanes traveling in opposite directions. Similar to neighbor relationships, the actors 
          need not be on immediately opposite lanes.
\end{enumerate}

%TODO: add figure

\subsection{Hierarchical Graph Construction Algorithm}
\label{subsec:construction_algorithm}

The actor graph construction follows a two-phase approach that separates relation discovery from 
graph building, enabling hierarchical processing and preventing redundant edge creation.

\subsubsection{Phase 1: Relation Discovery}
\label{subsubsec:discovery_phase}

The discovery phase identifies all potential relationships between actor pairs based on distance 
constraints and map graph structure. For each pair of actors $(A, B)$ at timestep $t$, the 
algorithm determines their primary lanes and checks if a path exists in the map graph connecting 
them. The path structure determines the relationship type:

\begin{itemize}
    \item \textbf{Following/Leading}: If both actors are on the same lane, or if the path consists 
          entirely of following edges.

    \item \textbf{Neighbor}: If the path contains exactly one neighbor edge and the remaining edges 
          are following edges. Note that this does not require the actors to be on immediately 
          adjacent lanes---the path may contain multiple following edges before and after the single 
          neighbor edge.

    \item \textbf{Opposite}: If the path contains exactly one opposite edge and the remaining edges 
          are following edges. Similar to neighbor relationships, this does not require the actors 
          to be on immediately opposite lanes.
\end{itemize}

Each relationship is validated using two distance checks: the lane-based path length ensures that 
curved roads are properly accounted for, while the Euclidean distance ensures that only actors in 
close proximity are considered. All discovered relationships are stored in a relations dictionary 
for the construction phase.

%TODO: add figure

\subsubsection{Phase 2: Hierarchical Graph Construction}
\label{subsubsec:construction_phase}

The construction phase builds the actor graph by adding edges in a hierarchical order, ensuring 
that higher-priority relationships are established first and redundant edges are prevented.

\paragraph{Construction Order}

Edges are added in three stages, following the semantic hierarchy:

\begin{enumerate}
    \item \textbf{Leading/Following edges} (highest priority): All following/leading relationships 
          are processed first, sorted by path length (shortest first).

    \item \textbf{Neighbor edges} (medium priority): Neighbor relationships are processed second, 
          sorted by absolute path length.

    \item \textbf{Opposite edges} (lowest priority): Opposite relationships are processed last, 
          sorted by absolute path length.
\end{enumerate}

\paragraph{Redundancy Prevention}

Before adding an edge between actors $A$ and $B$, the algorithm checks if a path already exists 
in the current actor graph with length $\leq$ the maximum node distance for that relation type. 
This prevents redundant edges that would create triangles or longer indirect paths when shorter 
direct paths exist.

For leading/following and neighbor relationships, each direction ($A \to B$ and $B \to A$) is 
checked independently, as these relationships can be asymmetric. For opposite relationships, if 
either direction has an existing path, both directions are rejected, as opposite relationships are 
inherently symmetric.

The path checking uses breadth-first search to find the shortest path in the actor graph, counting 
the number of edges (not nodes) in the path. This ensures that if a path exists with $k$ edges 
where $k \leq \text{max\_node\_distance}$, the direct edge is not added.

The graph is updated immediately after each edge addition, ensuring that path checks for 
subsequent edges use the current graph state. This incremental approach guarantees that the 
hierarchical construction works correctly and prevents race conditions.

%TODO: add figure

\subsubsection{Input Parameters}
\label{subsubsec:parameters}

Table~\ref{tab:construction_parameters} summarizes all input parameters for the actor graph 
construction algorithm, their descriptions, and the values used in our experiments.

\begin{table}[h]
\centering
\caption{Input parameters for actor graph construction}
\label{tab:construction_parameters}
\begin{tabular}{llp{6cm}l}
\hline
\textbf{Parameter} & \textbf{Type} & \textbf{Description} & \textbf{Value} \\
\hline
\multicolumn{4}{l}{\textit{Distance limits (discovery phase)}} \\
\hline
\texttt{max\_distance\_lead\_veh\_m} & float & Maximum distance in meters for leading/following 
relationships & 100 \\
\texttt{max\_distance\_neighbor\_forward\_m} & float & Maximum distance in meters for forward 
neighbor relationships & 50 \\
\texttt{max\_distance\_neighbor\_backward\_m} & float & Maximum distance in meters for backward 
neighbor relationships & 50 \\
\texttt{max\_distance\_opposite\_forward\_m} & float & Maximum distance in meters for forward 
opposite relationships & 100 \\
\texttt{max\_distance\_opposite\_backward\_m} & float & Maximum distance in meters for backward 
opposite relationships & 10 \\
\hline
\multicolumn{4}{l}{\textit{Node distance limits (construction phase)}} \\
\hline
\texttt{max\_node\_distance\_leading} & int & Maximum number of edges in actor graph path for 
leading/following & 3 \\
\texttt{max\_node\_distance\_neighbor} & int & Maximum number of edges in actor graph path for 
neighbor & 2 \\
\texttt{max\_node\_distance\_opposite} & int & Maximum number of edges in actor graph path for 
opposite & 2 \\
\hline
\multicolumn{4}{l}{\textit{Timestep configuration}} \\
\hline
\texttt{delta\_timestep\_s} & float & Time step increment in seconds for temporal graph 
creation & 1.0 \\
\hline
\end{tabular}
\end{table}

The distance limits in the discovery phase determine which potential relationships are identified, 
while the node distance limits in the construction phase control redundancy prevention. The 
asymmetric limits for opposite relationships (100m forward vs. 10m backward) reflect that vehicles 
traveling in opposite directions are more relevant when they are ahead rather than behind.

All parameters are configurable and can be adjusted based on the specific application 
requirements, road network characteristics, or desired graph density. The values shown in 
Table~\ref{tab:construction_parameters} represent default values that work well for urban traffic 
scenarios, but may need adjustment for highway scenarios (longer distances) or dense urban 
intersections (shorter distances).

%TODO: add examples

\subsection{Time based graph representations}
The graph creation strategy described above is assuming to create a graph for a single timestep. 
One possible extension is to create a graph which contains information for multiple timesteps. In 
order to create a time indexed graph, first the actor graphs for all timesteps are created. The 
joint set of all nodes across all timesteps is then determined. Node attributes of single timestep 
graphs are combined into a tensor of node attributes, introducing a new dimension for the time 
index. While the nodes are somehow time independent by this strategy, the edges are modeled as 
time dependent. Here it is utilized, that there can be multiple edges between the same pair of 
nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the 
time stamp of the edge.

This extension allows the graph to easily represent changing states, e.g. the relationship between 
two vehicles changes from neighbor to leading vehicle, i.e. a cut in happened. One important 
special case for this type of processing is to handle actors, which are not present in all time 
steps, i.e. they appear and disappear during the time window. Especially if node attributes are 
used in the processing, a suitable imputation strategy is necessary to avoid non-equal dimensions 
of tensors for node attributes.

The methods derived in the following chapters can be applied to this extended graph representation 
as well, even though the focus in this paper will be on single timestep graphs.

### 2025-11-23 article/sections/existing_approaches.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
 \section{Existing coverage and analysis approaches}
@@
 definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
+elements plus actors' self-representations; a situation is the set of all circumstances relevant for 
+behaviour decision at a given moment, derived from the scene but reflecting the actor's goals and 
 values; and a scenario is a temporal development of several scenes in sequence, involving 
@@
 measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-23 article/sections/graph_embeddings.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
 \section{Graph Embeddings for Traffic Scene Analysis}
@@
 and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-23 article/sections/introduction.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
 \section{Introduction}
@@
 followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-23 article/sections/summary.tex
article tex seperated. Graph construction part added

--- 2025-11-19 article/article.tex
@@
-\documentclass{article}
-\usepackage{cite}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{booktabs}
-\usepackage{svg}
-\usepackage{subcaption}
-
-\title{A Graph-based Framework for Coverage Analysis in Autonomous Driving \newline \newline Preprint}
-
-\author{Thomas Mühlenstädt and Marius Bause}
-\date{\today}
-
-\begin{document}
-
-\maketitle
-
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
 \section{Summary}
@@
 \end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{lit}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
-\end{document}

### 2025-11-25 article/article.tex
add article tex

--- 2025-11-19 article/article.tex
@@
 
-\section{Abstract}
-
-tbd
-
-% $$\hat{r} = \frac{N}{E}$$
-
-% $$LL_{\lambda} = \frac{1}{2}\chi^2_{0.025, 2N}$$
-% $$UL_{\lambda} = \frac{1}{2}\chi^2_{0.975, 2N + 2}$$
-
-% $$LL_{r} = LL_{\lambda} / E$$
-% $$UL_{r} = UL_{\lambda} / E$$
-
-\section{Introduction}
-
-In autonomous driving, coverage analysis is a crucial step to ensure the safety and reliability of the system.
-In most situations, coverage arguments are collected either per coverage factor, or maybe up to 2 or 3 factor interactions.
-See for example \cite{foretellix2019} for an production grade implementation of state of the art coverage analysis.
-In contrast to existing approaches, this paper proposes a graph-based framework for coverage analysis.
-There are already other graph-based approaches for analysing and representing traffic scenes.
-However, the work in that paper is not specifically focused on coverage analysis.
-Hence in this paper, graph based traffic scene representations are utilized for coverage analysis.
-
-This paper is structured as follows: In chapter \ref{chapter:existing_approaches}, existing coverage and analysis approaches are discussed.
-The basis for the following chapters is layed in chapter \ref{chapter:defining_a_traffic_scene_graph}. Afterwards, two coverage approaches 
-using the traffic scene graphs are discussed in chapter \ref{chapter:create_subgraphs_for_coverage_analysis} and 
-chapter \ref{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}.
-Finally, the developed methods are applied to Carla and Argoverse 2.0 data in chapter \ref{chapter:application} 
-followed by a short summary and outlook.
-
-\section{Existing coverage and analysis approaches}
-\label{chapter:existing_approaches}
-
-add Master thesis Johannes
-
-There exist a large amount of literature on coverage analysis in the context of autonomous driving.
-In the following, some of the most relevant approaches are discussed.
-
-The PEGASUS project (\cite{pegasus2019method}) introduces a systematic, scenario-based methodology for the verification and 
-validation of highly automated driving functions, addressing the impracticality of traditional 
-distance-based testing approaches. The core of the method is a six-layer model used to 
-systematically describe and structure the driving environment, encompassing factors from road geometry to weather conditions. 
-This framework is particularly relevant for coverage analysis, as it facilitates a structured decomposition of the vast, 
-continuous test space into discrete, manageable "logical scenarios". By systematically 
-parameterizing and exploring these logical scenarios across simulation, proving ground, and field tests, 
-the methodology aims to ensure the completeness of relevant test runs and provides a structured 
-foundation for arguing that the automated system has been adequately tested across its entire operational design 
-domain (ODD). This shift from random sampling to a structured, scenario-based approach allows 
-for a more efficient and comprehensive assessment to generate evidence for a final safety argumentation
-
-In a similar direction, \cite{deGelder2022ontology} focus on defining ontoligies for automated vehicles
-in an object oriented manner, with the intention to have a clearly defined and implementable
-structures for scenarios, delivering a clear linkage to coverage arguments.
-
-The authors of (\cite{Ries2021traj_clustering}) address the challenge of validating automated driving systems 
-in complex urban environments by proposing a trajectory-based clustering method for real-world driving data. 
-They extract motion trajectories and associated features from urban driving recordings, apply unsupervised 
-clustering to group similar driving behaviours/scenes, then analyze the resulting clusters to reveal common 
-scenario types and redundancies. Their approach enables structuring the enormous scenario space, 
-supporting more efficient test-set generation and scenario selection for automated vehicle verification. While 
-not explicitly focused on coverage analysis, their approach is a good starting point for coverage analysis.
-
-In \cite{Ulbrich2015scene}, the authors identify that key concepts in automated driving—namely scene, 
-situation, and scenario—are inconsistently defined in the literature, which complicates the 
-development, testing and validation of driving-automation modules. They propose clear 
-definitions: a scene is a snapshot of the environment including dynamic and static 
-elements plus actors’ self-representations; a situation is the set of all circumstances relevant for 
-behaviour decision at a given moment, derived from the scene but reflecting the actor’s goals and 
-values; and a scenario is a temporal development of several scenes in sequence, involving 
-actions/events and goals/values. The paper also provides example implementations of these 
-definitions in the context of automated-vehicle systems and how they interface with perception, 
-planning and control, and testing.
-
-Another important reference is the SAE International recommended practice \cite{ORAD2021taxonomy}, which 
-provides many foundational terms in the context of autonomous driving.
-It provides a functional taxonomy and clear definitions for key terms related to 
-driving automation systems (DAS) used in on-road motor vehicles. The document defines six 
-levels of driving automation (Levels 0 through 5) based on the role of three primary actors — 
-the human driver, the driving automation system (DAS), and other vehicle systems/components. It introduces 
-important related terms such as the Dynamic Driving Task (DDT), Operational Design 
-Domain (ODD), Automated Driving System (ADS), and Vehicle Motion Control, 
-among others. The 2021-04 revision (superseding the 2018 version) was developed in 
-collaboration with ISO/TC 204/WG14 to harmonise global terminology and improve clarity for multi-discipline audiences 
-(engineering, legal, media). The document emphasises that it is descriptive (not normative); 
-it does not prescribe specifications or impose performance requirements for DAS.
-
-The paper \cite{DBLP:journals/corr/abs-1801-08598} proposes a scenario-based framework for developing and 
-validating automated-driving systems across different development phases. It 
-introduces three abstraction levels of scenarios—functional, logical, and 
-concrete—and discusses how these can be transformed for use in testing. The authors note that 
-existing parameter-selection methods, such as equivalence-class or combinatorial testing, lack a 
-systematic way to determine meaningful test coverage, highlighting coverage analysis as 
-an open challenge in scenario-based validation.
-
-In \cite{Ammann_Offutt_2008}, a standard textbook on software testing although not specifically focused on 
-autonomous driving, the authors provide an introduction to software testing, which is a 
-good starting point for coverage analysis. They introduce the concept of coverage and discuss the 
-different types of coverage, such as statement coverage, branch coverage, condition coverage, 
-and decision coverage. They also discuss the different techniques for measuring coverage, such as 
-branch coverage, condition coverage, and decision coverage.
-
-A much more focussed example of coverage analysis in the context of autonomous driving is provided 
-by \cite{foretellix2019} in a blog 
-post, discussing topics like items, parameters and 
-coverage buckets and performance metrics like time to collision and error collections.
-
-The authors of (\cite{wachenfeld2016release}) investigate the significant challenge of safety validation and 
-production release for fully autonomous vehicles, positing that established testing concepts are insufficient 
-as they fundamentally rely on the human driver's ability to intervene as a safety backup. 
-The authors introduce the "approval-trap," a statistical argument demonstrating the unfeasibility of proving 
-superior safety through real-world driving, which would necessitate billions of test kilometers. 
-This validation gap is presented as a fundamental problem of coverage analysis, where the
-technical system must now demonstrably cover the vast operational domain previously managed by the 
-human. The paper concludes that overcoming this challenge requires a paradigm shift toward new 
-test case generation methodologies, such as critical scenario identification, and the extensive
- use of validated simulation-based tools to achieve sufficient test coverage efficiently.
-
-Well known standards for coverage analysis in the context of autonomous driving are the 
-ISO 21448 (\cite{iso21448}) and the UL 4600 (\cite{ul4600}).
-The ISO 21448 is a standard for the safety of the intended functionality of road vehicles, 
-including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and 
-the measurement of coverage.
-The UL 4600 is a standard for the safety of autonomous products, including automated driving systems.
-It defines a framework for coverage analysis, including the definition of coverage criteria and the 
-measurement of coverage.
-
-\section{Defining a traffic scene graph}
-\label{chapter:defining_a_traffic_scene_graph}
-
-still to be added: \url{https://sagroups.ieee.org/adwg/wp-content/uploads/sites/661/2024/10/ADWG_STV2_whitepaper.pdf}
-
-still to be added: \url{https://www.vvm-projekt.de/securedl/sdl-eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJpYXQiOjE3NDc2OTIxNjgsImV4cCI6MTc0Nzc4MjE2OCwidXNlciI6MCwiZ3JvdXBzIjpbMCwtMV0sImZpbGUiOiJmaWxlYWRtaW4vdXNlcl91cGxvYWQvVlZNZXRob2Rlbl9adXNhbW1lbmZhc3NlbmRlcl9BYnNjaGx1c3NiZXJpY2h0X3YxLjBfenVyX1Zlc_f9d402f084002b57/VVMethoden_Zusammenfassender_Abschlussbericht_v1.0_zur_Veroeffentlichung.pdf}
-
-\cite{newman2010networks}
-
-\cite{bagschik2018ontologybasedscenecreation}
-
-\cite{fremont2020scenic}
-
-\subsection{Time based graph representations}
-The graph creation strategy described above is assuming to create a graph for a single timestep. One possible 
-extension is to create a graph which contains information for multiple timesteps. In order to create a time indexed graph,
-first the actor graphs for all timesteps are created. The joint set of all nodes across all timesteps is then determined.
-Node attributes of single timestep graphs are combined into a tensor of node attributes, introducing a new dimension for the time index.
-While the nodes are somehow time independent by this strategy, the edges are modeled as time dependent. Here it is utilized, that 
-there can be multiple edges between the same pair of nodes, keeping potential edge attributes and adding an additional edge attribute to indicate the
-time stamp of the edge.
-
-This extension allows the graph to easily represent changing states, e.g. the relationship between two vehicles 
-changes from neighbor to leading vehicle, i.e. a cut in happened. One important special case for this type of processing is 
-to handle actors, which are not present in all time steps, i.e. they appear and disappear during the time window. Especially 
-if node attributes are used in the processing, a suitable imputation strategy is necessary to avoid 
-non-equal dimensions of tensors for node attributes.
-
-The methods derived in the following chapters can be applied to this extended graph representation as well, even though 
-the focus in this paper will be on single timestep graphs.
-
-\section{Analysing a traffic scene with a graph}
-
-Having defined a graph-based traffic scene representation, we can now analyse the coverage of the system.
-Two methodologies are proposed for this purpose:
-One is to define archetypes of traffice scenes, and to compare graphs from observed traffic scenes to these archetypes.
-The second one is to translate graphs to graph embeddings, and then to compare the embeddings of different sets of traffic scenes.
-
-\section{Create subgraphs for coverage analysis}
-\label{chapter:create_subgraphs_for_coverage_analysis}
-
-There is a lot of knowledge in the literature on how to define archetypes of traffic scenes.
-Once an archetype is defined, a special property of graphs can be used. 
-Two graphs are isomorphic if they have the same structure, regardless of the node and edge labels.
-As the archetypes are not necessarily involving a lot of actors, these are more like subsets of actual traffic scenes.
-A very simple example might be 2 vehicles on the same lane, driving in the same direction and another vehicle driving on a neighboring lane.
-This situation can be represented by a graph with 3 nodes and 2 edges. 
-In most real traffic situations however, there will be additional actors present, so that we are not searching for isomorphic graphs, but rather want to check if any subgraph of $G$ is isomorphic to the archetype graph $A$.
-This is an example of a subgraph isomorphism problem.
-While this problem is NP-hard, the graphs considered here are rather small, so the computational time is reasonable.
-One such algorithm is the VF2 algorithm, which is implemented in the NetworkX library (see \cite{cordella2004subgraph}).
-The strategy we are then applying is the following:
-
-\begin{enumerate}
-    \item Define a set of subgraphs $S$ that are considered to be archetypes of traffic scenes, e.g. unprotected left turns with opposite traffic or lead vehicle following situations.
-    \item Define which node and edge attributes are considered for the isomorphism check.
-    \item Create an empty dataframe $C$ with a column for each subgraph in $S$
-    \item Define the set of traffice scenes (e.g. from Carla or Argoverse) defined as graphs $G$
-    \item For each graph $G$, check if any subgraph of $G$ is isomorphic to any subgraph in $S$ and note the result in a new row in table $C$
-\end{enumerate}
-
-This strategy can be described to some degree as a bottom up approach: Starting from a detail level, individual situations are defined.
-Then going upwards to different datasets, it is checked, if the archetype is present.
-Also, follow up analysis of the created coverage dataframe can be performed. For example,
-
-\begin{itemize}
-    \item The distribution of numeric attributes like speed and and distance to other actors can be visualized for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-    \item It can be cross tabulated, which combinations of archetypes are jointly present in a traffic scene.
-    \item Pass Fail rates or other AV performance metrics can be calculated for the subset of all traffic scenes which are subgraph isomorphic to an archetype.
-\end{itemize}
-
-In the above approach, the archetypes are defined manually. An alternative approach is to automatically extract all possible archetypes 
-from a dataset of real world traffic observation data.
-
-xxxxxx
-
-TODO MARIUS: Describe the approach of automatically extracting archetypes from real world traffic observation data.
-
-xxxxxx
-
-\section{Graph Embeddings for Traffic Scene Analysis}
-\label{chapter:implementation_of_graph_embeddings_for_traffic_scene_analysis}
-
-This section describes the concepts, implementation and application of graph embeddings to traffic scene graphs. 
-The implementation follows a comprehensive approach to learning graph representations through self-supervised 
-contrastive learning.
-
-Embeddings are a widely used method to translate raw data like images or text into an embedding space in order to be 
-able to perform machine learning tasks on them. One well known example of this is the Word2Vec model, which is used to 
-translate words into a 
-vector space, where the distance between vectors can be used to measure the similarity between 
-words (\cite{mikolov2013efficientestimationwordrepresentations}).
-
-In the context of traffic scene graphs, embeddings are used to translate the graph structure into a vector space, 
-where the distance between vectors can be used to measure the similarity between traffic scenes.
-This is useful for coverage analysis, as it allows to compare traffic scenes among each other.
-For example, two traffic scenes can be considered similar if the distance between their embeddings is small. This enables
-to search for a most similar simulation scenario given a real world scenario, to identify areas with near duplicates or 
-to easily visualize structures in the embedding space, which in the original space of all possible traffic scenes
-would not be possible.
-
-Graph neural networks (GNNs) are a class of neural networks that are designed to process graph-structured data and have 
-gained a lot of popularity in the last years, see for example (add references).
-
-In this paper, a network architecture using a Graph Isomorphism Network with Edge features (GINE) as described 
-in \cite{hu2020strategiespretraininggraphneural} is used to generate embeddings for traffic scene graphs as implemented in the 
-pytorch geometric library (\cite{Fey/Lenssen/2019}). Main reason for using this specific architecture is that is capable of
-learning embedding representations not only on the graph structure itself but on both node and edge attributes. Other 
-network architectures like GraphSAGE or GAT are not capable of this. (TODO: check if this is true)
-
-\begin{figure}[h]
-    \centering
-    %\includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture.pdf}
-    \includegraphics[width=0.8\textwidth]{plots/graph_gine_architecture_manually_updated.png}
-    \caption{Model architecture for the Graph Isomorphism Network with Edge features (GINE).}
-    \label{fig:graph_gine_architecture}
-\end{figure}
-
-The exact architecture of the model is shown in Figure \ref{fig:graph_gine_architecture}. The features 
-used are the actor type (as a one hot encoding),
-the actor speed (float), if the actor is on an intersection (boolean) and if the actor changed its 
-lane since the last timestep (boolean) for the nodes.
-For the edges, the edge type (as a one hot encoding) and the path length (float) between the two nodes are used.
-
-The model has been trained on the CARLA and Argoverse 2.0 datasets using self-supervised contrastive learning.
-
-Training employed a self-supervised contrastive objective on mini-batches of 128 graphs. 
-For each batch, two correlated views of every graph were created by 
-perturbing continuous attributes with zero-mean Gaussian noise ($\sigma=0.1$) applied to node 
-longitudinal speed and edge path length. A four-layer GINE encoder 
-(hidden width 96) produced graph-level representations via the concatenation of 
-mean, max, and sum pooling, followed by an embedding MLP (yielding 256-dimensional embeddings) and a projection head.
-The contrastive loss was a temperature-scaled cross-entropy over in-batch similarities (cosine similarity of $\ell_2$-normalized projections, temperature $\tau=0.1$), maximizing agreement of the two views of the same graph while contrasting against other graphs in the batch.
-Optimization used Adam with an exponentially decaying learning rate, initialized at 0.02 and multiplied by 0.75 across successive stages.
-This setup follows established practice in contrastive pretraining for GNNs \cite{hu2020strategiespretraininggraphneural} 
-and is implemented using PyTorch Geometric \cite{Fey/Lenssen/2019}.
-
-\section{Application}
-\label{chapter:application}
-
-\subsection{Argoverse 2.0}
-
-\cite{Argoverse2}
-
-TODO MARIUS: Describe the Argoverse 2.0 dataset.
-
-\subsection{Carla}
-
-CARLA (Car Learning to Act, \cite{Dosovitskiy17Carla}) is an open-source simulator specifically designed for autonomous driving research and 
-development. It provides a highly realistic urban driving environment with 
-diverse road layouts, weather conditions, and traffic scenarios. The simulator features a comprehensive 
-sensor suite simulation, flexible API for scenario creation, and supports both learning-based 
-and traditional autonomous driving approaches. CARLA enables researchers to test and 
-validate autonomous vehicle systems in a safe, controllable environment before real-world deployment.
-
-The simulator has gained widespread adoption across both academic and industrial settings. In research, CARLA serves as a standard platform for developing and 
-benchmarking autonomous driving algorithms, including reinforcement learning approaches for vehicle control and sensor fusion 
-techniques \cite{codevilla2019exploringlimitationsbehaviorcloning}. Industry applications include 
-virtual testing of production autonomous vehicle systems, scenario-based validation pipelines, and integration 
-with hardware-in-the-loop testing frameworks \cite{jaeger2023hiddenbiasesendtoenddriving}. CARLA 
-is also extensively used in autonomous driving competitions and challenges, providing a common evaluation 
-environment for comparing different approaches across research groups worldwide.
-
-Here, Carla version $0.9.15$ is used. The CARLA version $0.10.0$ is not used, because it had only 2 maps and
-Mine\_1 (which is not really normal roads) at the start of this project.
-Specifically, the following maps were used: Town01, Town02, Town03, Town04, Town05 and Town07. Plots
-of these maps are shown in Figure \ref{fig:carla_maps}.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.8\textwidth]{plots/carla_maps_used.png}
-\caption{Overview of CARLA maps used in the simulation study: Town01, Town02, Town03, Town04, Town05, and Town07.
-These maps provide diverse urban driving environments with varying road layouts, intersections, and traffic patterns.}
-\label{fig:carla_maps}
-\end{figure}
-
-The data generation script implements sophisticated behavior control mechanisms to create 
-diverse and realistic traffic scenarios. Multiple vehicle types including trucks, motorcycles, 
-and regular cars are spawned with varying probabilities, each exhibiting different behavioral
-characteristics such as speed preferences, following distances, and lane-changing tendencies. 
-The script incorporates dynamic behavior modifications during simulation, including random slowdowns, 
-periodic behavior changes, and adaptive responses to traffic conditions, resulting in rich and 
-varied traffic scene data across multiple CARLA maps and simulation iterations. The simulation runs
-have between 20 and 60 vehicles each.
-
-The resulting data consists of xxx scenes with 11 seconds of simulation time each, in order to have a 
-similar data size as the Argoverse 2.0 dataset.
-
-\subsection{Subgraph Isomorphism Coverage Analysis}
-
-We apply the subgraph isomorphism approach to both the  CARLA and the Argoverse data to identify coverage scenarios.
-The archetypes used here are defined manually and described in table xxx.
-The chosen archetypes are typical traffic situations like e.g.
-lead vehicle situations, lane change situations or different combinations of opposite direction vehicles.
-Also, information like which traffic actors are on an intersection are incorporated into the archetypes.
-The results are shown in Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} to Figure \ref{fig:combined_distributions_plots_speed_distance_argo}.
-In Figure \ref{fig:subgraph_isomorphism_coverage_barcharts} there is a clear signal that for both datasets the coverage is not uniform per the different archetypes.
-Also, the distribution for the Carla dataset is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Even worse, the Carla dataset is nearly completely missing out e.g. on the cut\_out\_intersection archetype, clearly indicating a coverage gap in the Carla dataset.
-A next step of analysis is to check, which archetypes are occuring simultaneously in a traffic scene. 
-Figure \ref{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios} shows the agreement matrix for the manually defined coverage scenarios for CARLA and Argoverse.
-The heatmaps show the percentage of agreement between the manually defined coverage scenarios for CARLA and Argoverse.
-
-TODO: Write a sentence about the agreement matrices once using the actual actor graph defintion.
-
-A last example of how to use the assignment of archetypes to traffic scenes is to check the parameter distribution for the speed and path length of the traffic scenes.
-Figure \ref{fig:combined_distributions_plots_speed_distance_carla} and Figure \ref{fig:combined_distributions_plots_speed_distance_argo} show the parameter distribution for the speed and path length of the traffic scenes for CARLA and Argoverse.
-The plots show that the distribution for the CARLA data is not close to the Argoverse distribution, indicating that the CARLA data is not representative for the Argoverse data.
-Given the assignment of archetypes to traffic scenes based on the actors can be used in further analysis not done here, e.g. to check more divers parameters and distributions.
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_carla_coverage_barchart.png}
-        \caption{CARLA coverage scenarios}
-        \label{fig:subgraph_isomorphism_carla_coverage_barchart}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_argo_coverage_barchart.png}
-        \caption{Argoverse coverage scenarios}
-        \label{fig:subgraph_isomorphism_argo_coverage_barchart}
-    \end{subfigure}
-    \caption{Coverage barcharts for the manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_coverage_barcharts}
-\end{figure}
-
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[width=0.8\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-%     \caption{Agreement matrix for manually defined coverage scenarios for CARLA.}
-%     \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-% \end{figure}
-
-\begin{figure}[h]
-    \centering
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_carla.png}
-        \caption{Agreement matrix for CARLA.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_carla}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}[b]{0.48\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{plots/subgraph_isomorphism_agreement_matrix_manual_scenarios_argo.png}
-        \caption{Agreement matrix for Argoverse.}  
-        \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios_argo}
-    \end{subfigure}
-    \caption{Agreement matrix for manually defined coverage scenarios for CARLA and Argoverse.}
-    \label{fig:subgraph_isomorphism_agreement_matrix_manual_scenarios}
-\end{figure}
-
-% \begin{table}[htbp]
-%     \caption{My Table}
-%     \label{tab:mytable}
-%     \begin{tabular}{rrrrr}
-%     \toprule
-%     scn 1 & scn 2 & scn 3 & scn 4 & size \\
-%     \midrule
-%     False & False & False & False & 613 \\
-%     False & False & True & False & 3 \\
-%     True & True & False & True & 377 \\
-%     True & True & True & True & 7 \\
-%     \bottomrule
-%     \end{tabular}
-% \end{table}
-% % \cite{goodfellow2016deep} % needed?
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_carla.png}
-    \caption{Parameter distribution for speed and path length for CARLA.}
-    \label{fig:combined_distributions_plots_speed_distance_carla}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.85\textwidth]{plots/combined_distributions_plots_speed_distance_argo.png}
-    \caption{Parameter distribution for speed and path length for Argoverse.}
-    \label{fig:combined_distributions_plots_speed_distance_argo}
-\end{figure}
-
-\subsection{Graph Embeddings Coverage Analysis}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/train_test_graph_embeddings_loss_plot.png}
-    \caption{Training and test loss for the graph embeddings model trained jointly on CARLA and Argoverse 2.0 data.}
-    \label{fig:train_test_graph_embeddings_loss_plot}
-\end{figure}
-
-The resulting embeddings have been analysed in a number of ways. For plausibility checks, for a number of 
-randomly samples scenarios, the scenario with 
-closest embedding vector (euclidean distance) has been visualized. 
-This is shown in Figure \ref{fig:plausibility_checks}. This 
-includes comparison between CARLA and Argoverse scenarios.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\textwidth]{plots/graph_embeddings_pca_tsne_test_plot.png}
-    \caption{PCA and t-SNE visualization of the embedding space for Carla and Argoverse 2.0 scenarios.}
-    \label{fig:embedding_space}
-\end{figure}
-
-As a next step, the embedding space has been analysed using PCA and t-SNE. This is shown in Figure \ref{fig:embedding_space}. This can be done in a number of 
-different flavors, like for example:
-
-\begin{itemize}
-    \item distinguishing between CARLA and Argoverse scenarios by a color coding, 
-    \item visualiizing just the CARLA scenarios and color code them by the map, in order to see if the maps deliver different types of scenarios,
-    \item by calculating a density distribution of the embedding space for the Argoverse scenarios, and to check if for a regions with a density 
-    about a certain threshold, there exist CARLA scenarios, i.e. do a coverage analysis in the embedded space
-    \item do the same vice versa, i.e. to check for relevance of CARLA scenarios.
-\end{itemize}
-
-As the embedding space is a normal metric space, representing scenarios across a large range of different driving situations, these embeddings can be used also for 
-many other tasks, like for example clustering, anomaly detection, similarity search, etc.
-
-And, as the main task considered here is coverage analysis, the embeddings can be used to check if a target distribution is met by a test distribution in the embedded space. 
-Specifically, considering the Argoverse 2.0 scenarios as the target distribution, and the CARLA scenarios as the test distribution, the embeddings can be used to check 
-if the CARLA scenarios cover the Argoverse 2.0 scenarios in the embedded space.
-
-\section{Summary}
-
-\begin{itemize}
-    \item A method for using traffic scenes represented as graphs for coverage analysis has been developed in this paper.
-    \item Starting from defining rules for the actor graph construction, the resulting graphs are then used by 2 different approaches for coverage analysis.
-    \item Firstly, a non-AI approach is used where a set of archetypes is defined (which is done here manually but in principle could be done automatically as well).
-    \item Then it is checked, which archetypes are present in a traffic scene and how often.
-    \item Secondly, a graph embedding approach is used to create a vector space of the traffic scenes and to check if the target distribution is met by the test distribution in the embedded space.
-    \item The results for this approach are shown in Section \ref{chapter:application}, being applied to Carla and Argoverse 2.0 data.
-    \item The main advantage of this approach over traditional coverage analysis approaches is that it very easily scales to a large number of traffic scenarios.
-    \item In other approaches (cite TNO Streetwise) it is necessary to manually define for each scenario type how it is defined.
-Here the only thing necessary is to define the actor graph construction rules as well as potentially the archetypes as graphs.
-    \item 
-    \item 
-\end{itemize}
+\input{sections/abstract}
+\input{sections/introduction}
+\input{sections/existing_approaches}
+\input{sections/defining_traffic_scene_graph}
+\input{sections/analysing_traffic_scene}
+\input{sections/create_subgraphs}
+\input{sections/graph_embeddings}
+\input{sections/application}
+\input{sections/summary}
 
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_0.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_1.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_2.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_3.png}
-
-    \includegraphics[width=0.95\textwidth]{plots/graph_embeddings_comparison_plausibility_check_4.png}
-    \caption{Plausibility checks showing graph embedding comparisons.
-For randomly sampled scenarios, the most similar scenarios based on embedding distance (both Euclidean and cosine) are visualized, including comparisons between CARLA and Argoverse scenarios.}
-    \label{fig:plausibility_checks}
-\end{figure}
-
 \end{document}

### 2025-11-27 scripts/carla_actor_graph_creation.py
graph creation scripts

@@ -1,0 +1,99 @@
+import pandas as pd
+from pathlib import Path
+import matplotlib.pyplot as plt
+import seaborn as sns
+import os
+from scipy.spatial import distance
+import xml.etree.ElementTree as ET
+from shapely.geometry import Polygon, LineString
+import pandas as pd
+#from src.generate_traffic_data import clean_carla, spawn_scene, run_scene
+import carla
+import os
+os.getcwd()
+from datetime import datetime
+import time
+import random
+from tqdm import tqdm
+import networkx as nx
+import numpy as np
+import pickle
+import json
+#os.chdir('../..')
+#os.getcwd()
+
+from graph_creator.MapGraph import MapGraph
+from graph_creator.ActorGraph import ActorGraph
+from graph_creator.utilities import make_node_edge_df, check_offroad_actors
+from graph_creator.plot_graphs import plot_lane_map_advanced, add_actors_to_map, add_actor_edges_to_map
+
+
+files = os.listdir("/home/tmuehlen/repos/graph_coverage/carla/data")
+#scn_ids = [file.split("_")[1] for file in files if "tracks_with_min_distance_to_lane" in file]
+scn_ids = [file.split("scene_")[1].split("_tracks")[0] for file in files if "tracks.parquet" in file]
+scn_ids = sorted(scn_ids)
+
+#scn_ids = [scn_id for scn_id in scn_ids if scn_id > "2025-09-05"]
+print(len(scn_ids))
+#scn_ids
+
+# read the actor graph setting from the json file
+from pathlib import Path
+# repo_root = Path("/Users/marius/code/graph_coverage")
+repo_root = Path("/home/tmuehlen/repos/graph_coverage")
+
+settings_name = "actor_graph_setting_1_50_50_10_20_20_4_4_4"
+with open(f"graph_settings/{settings_name}.json", "rb") as f:
+    actor_graph_setting = json.load(f)
+
+name_aug = settings_name
+os.makedirs(repo_root / "actor_graphs" / f"carla_{name_aug}", exist_ok=True)
+os.makedirs(repo_root / "actor_graphs" / f"carla_{name_aug}_components_nx", exist_ok=True)
+os.makedirs(repo_root / "actor_graphs" / f"carla_{name_aug}_nx", exist_ok=True)
+
+
+# run through all scens so far and create the graphs:
+#os.makedirs("/home/tmuehlen/repos/graph_coverage/actor_graphs/carla_w_intersection", exist_ok=True)
[…]

### 2025-11-28 scripts/carla_actor_graph_creation.py
create more data and scripts

@@ -25,7 +25,7 @@ import json
 from graph_creator.MapGraph import MapGraph
 from graph_creator.ActorGraph import ActorGraph
 from graph_creator.utilities import make_node_edge_df, check_offroad_actors
-from graph_creator.plot_graphs import plot_lane_map_advanced, add_actors_to_map, add_actor_edges_to_map
+#from graph_creator.plot_graphs import plot_lane_map_advanced, add_actors_to_map, add_actor_edges_to_map
 
 
 files = os.listdir("/home/tmuehlen/repos/graph_coverage/carla/data")
@@ -54,8 +54,14 @@ os.makedirs(repo_root / "actor_graphs" / f"carla_{name_aug}_nx", exist_ok=True)
 
 # run through all scens so far and create the graphs:
 #os.makedirs("/home/tmuehlen/repos/graph_coverage/actor_graphs/carla_w_intersection", exist_ok=True)
+
 count = 0
-for scn_id in tqdm(scn_ids):
+
+for i, scn_id in enumerate(scn_ids):
+    print(i, "/", len(scn_ids), scn_id)
+    if os.path.exists(repo_root / "actor_graphs" / f"carla_{name_aug}" / f"{scn_id}_map_graph.pkl"):
+        continue
+    
     g_map = MapGraph()
     g_map.read_graph_from_file(f'/home/tmuehlen/repos/graph_coverage/carla/data/scene_{scn_id}_map_graph.pickle')
     # tracks = pd.read_parquet(f'/home/tmuehlen/repos/graph_coverage/carla/data/scene_{scn_id}_tracks_with_min_distance_to_lane.parquet')
@@ -91,7 +97,7 @@ for scn_id in tqdm(scn_ids):
                 save_path = repo_root / "actor_graphs" / f"carla_{name_aug}_components_nx" / f"graph_{scn_id}_{timestamp_idx}_{component_idx}.pkl"
                 with open(save_path, "wb") as file:
                     pickle.dump(ag_carla.actor_components[ag_timestamps[timestamp_idx]][component_idx], file)
-        save_path = repo_root / "actor_graphs" / f"carla_{name_aug}_nx" / f"graph_{scn_id}_{str(timestamp).replace(".", "_")}.pkl"
+        save_path = repo_root / "actor_graphs" / f"carla_{name_aug}_nx" / f'graph_{scn_id}_{str(timestamp_idx).replace(".", "_")}.pkl'
[…]

### 2025-11-28 scripts/graph_embedding_training.py
create training config

@@ -4,9 +4,14 @@ import sys
 import pickle
 import torch
 import numpy as np
+import json
+import argparse
 from tqdm import tqdm
+from loguru import logger
 from sklearn.model_selection import train_test_split
 from torch.utils.data import Subset
+
+# python scripts/graph_embedding_training.py --config configs/training_config.json
 
 # Try to import DataLoader from loader (newer PyG) or data (older PyG)
 try:
@@ -31,50 +36,64 @@ def ensure_dir(path):
     if not os.path.exists(path):
         os.makedirs(path)
 
-def train():
+def train(config):
+    # Logger setup
+    log_file = config['training'].get('log_file', 'training.log')
+    # Configure logger to write to file
+    logger.add(log_file, rotation="10 MB")
+
     # Paths
-    carla_pattern = os.path.join(PROJECT_ROOT, "actor_graphs/carla_actor_graph_setting_1_50_50_10_20_20_4_4_4_components_nx/*.pkl")
-    argoverse_pattern = os.path.join(PROJECT_ROOT, "actor_graphs/argoverse_actor_graph_setting_1_50_50_10_20_20_4_4_4_components_nx/*.pkl")
+    carla_pattern = os.path.join(PROJECT_ROOT, config['data']['carla_pattern'])
+    argoverse_pattern = os.path.join(PROJECT_ROOT, config['data']['argoverse_pattern'])
     
-    print("Looking for graphs...")
+    logger.info("Looking for graphs...")
     graph_paths = glob.glob(carla_pattern)
     argoverse_graph_paths = glob.glob(argoverse_pattern)
     graph_paths.extend(argoverse_graph_paths)
     
-    print(f"Found {len(graph_paths)} graphs")
+    logger.info(f"Found {len(graph_paths)} graphs")
 
     # Checkpoints
-    checkpoint_dir = os.path.join(PROJECT_ROOT, "checkpoints")
+    checkpoint_dir = os.path.join(PROJECT_ROOT, config['training']['checkpoint_dir'])
     ensure_dir(checkpoint_dir)
 
     # Device
     device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
-    # device = "cpu"
-    print(f"Using device: {device}")
+    logger.info(f"Using device: {device}")
 
     # Data Loading
     graph_ds = GraphDataset(graph_paths)
     indices = np.arange(len(graph_ds))
[…]

### 2025-11-30 scripts/argoverse_actor_graph_creation.py
updates to embeddings

@@ -38,7 +38,7 @@ print(f"Found {len(scenario_folders)} scenarios")
 
 # read the actor graph setting from the json file
 settings_name = "actor_graph_setting_1_50_50_10_20_20_4_4_4"
-with open(f"graph_settings/{settings_name}.json", "rb") as f:
+with open(f"configs/graph_settings/{settings_name}.json", "rb") as f:
     actor_graph_setting = json.load(f)
 
 name_aug = settings_name

### 2025-11-30 scripts/carla_actor_graph_creation.py

@@ -43,7 +43,7 @@ from pathlib import Path
 repo_root = Path("/home/tmuehlen/repos/graph_coverage")
 
 settings_name = "actor_graph_setting_1_50_50_10_20_20_4_4_4"
-with open(f"graph_settings/{settings_name}.json", "rb") as f:
+with open(f"configs/graph_settings/{settings_name}.json", "rb") as f:
     actor_graph_setting = json.load(f)
 
 name_aug = settings_name

### 2025-11-30 scripts/graph_embedding_training.py

@@ -6,12 +6,13 @@ import torch
 import numpy as np
 import json
 import argparse
+from datetime import datetime
 from tqdm import tqdm
 from loguru import logger
 from sklearn.model_selection import train_test_split
 from torch.utils.data import Subset
 
-# python scripts/graph_embedding_training.py --config configs/training_config.json
+# python scripts/graph_embedding_training.py --config configs/embeddings/training_config_default.json
 
 # Try to import DataLoader from loader (newer PyG) or data (older PyG)
 try:
@@ -37,10 +38,24 @@ def ensure_dir(path):
         os.makedirs(path)
 
 def train(config):
+    # Setup timestamp and directories
+    timestamp = datetime.now().strftime("%Y%m%d_%H%M%S")
+    base_checkpoint_dir = os.path.join(PROJECT_ROOT, config['training'].get('checkpoint_dir', 'checkpoints'))
+    run_dir = os.path.join(base_checkpoint_dir, timestamp)
+    ensure_dir(run_dir)
+
     # Logger setup
-    log_file = config['training'].get('log_file', 'training.log')
+    log_file = os.path.join(run_dir, 'training.log')
     # Configure logger to write to file
     logger.add(log_file, rotation="10 MB")
+    logger.info(f"Starting training run: {timestamp}")
+    logger.info(f"Run directory: {run_dir}")
+
+    # Save config copy
+    config_save_path = os.path.join(run_dir, "config.json")
+    with open(config_save_path, 'w') as f:
+        json.dump(config, f, indent=4)
+    logger.info(f"Saved config to {config_save_path}")
 
     # Paths
     carla_pattern = os.path.join(PROJECT_ROOT, config['data']['carla_pattern'])
@@ -52,10 +67,6 @@ def train(config):
     graph_paths.extend(argoverse_graph_paths)
     
     logger.info(f"Found {len(graph_paths)} graphs")
-
-    # Checkpoints
-    checkpoint_dir = os.path.join(PROJECT_ROOT, config['training']['checkpoint_dir'])
-    ensure_dir(checkpoint_dir)
 
     # Device
     device = torch.device('cuda' if torch.cuda.is_available() else 'cpu')
@@ -164,7 +175,7 @@ def train(config):
             logger.info(f"Current Losses - Train: {total_losses['train'][-1]:.4f}, Test: {total_losses['test'][-1]:.4f}")
 
             # Checkpointing
[…]

### 2025-12-07 subgraphs/SubgraphIsomorphismChecker.py
intermediate subrgaph extract

@@ -1,0 +1,129 @@
+import pandas as pd
+import networkx as nx
+from tqdm import tqdm
+import pickle
+from typing import List, Dict, Optional
+from collections import defaultdict
+
+
+class IsomorphicGrapCoverageCounter:
+    def __init__(self, coverage_graphs: Dict[str, nx.DiGraph], graph_paths: List[str], node_match: List[str], edge_match: List[str]):
+        """
+        Initialize the IsomorphicGrapCoverageCounter.
+
+        Args:
+            coverage_graphs: Dictionary of coverage graphs.
+            graph_paths: List of graph paths.
+            node_match: List of node match attributes.
+            edge_match: List of edge match attributes.
+        """
+        self.coverage_graphs = coverage_graphs
+        self.graph_paths = graph_paths
+        self.node_match = node_match
+        self.edge_match = edge_match
+        self.cov_data = {}
+        for scen_name in self.coverage_graphs.keys():
+            self.cov_data[scen_name] = []
+
+        self.cov_data["degree"] = []
+        self.cov_data["density"] = []
+        self.cov_data["diameter"] = []
+        self.cov_data["path"] = []
+
+
+    def count_isomorphic_graphs(self):
+        for graph_path in tqdm(self.graph_paths, desc="Checking isomorphic graphs"):
+            with open(graph_path, "rb") as file:
+                ag_nx = pickle.load(file)
+
+            for key in self.coverage_graphs.keys():
+                GM = nx.algorithms.isomorphism.DiGraphMatcher(
+                    ag_nx, self.coverage_graphs[key],
+                    node_match=nx.algorithms.isomorphism.categorical_node_match(self.node_match, [None] * len(self.node_match)),
+                    edge_match=nx.algorithms.isomorphism.categorical_edge_match(self.edge_match, [None] * len(self.edge_match))
+                )
+                self.cov_data[key].append(GM.subgraph_is_isomorphic())
+            
+            self.cov_data["degree"].append(sum(dict(ag_nx.degree()).values()) / len(ag_nx.nodes()))
+            self.cov_data["density"].append(nx.density(ag_nx))
+            self.cov_data["diameter"].append(nx.diameter(ag_nx))
+            self.cov_data["path"].append(graph_path)
+
+        self.cov_data_df = pd.DataFrame(self.cov_data)
+
+    def count_node_coverage(self, max_graphs: int = 10000) -> pd.DataFrame:
+        """
+        Count how many times each node in each graph is covered by each subgraph type.
[…]

### 2025-12-07 subgraphs/subgraph_types.py

@@ -1,0 +1,169 @@
+"""
+Manually defined subgraph types for coverage analysis.
+
+This module contains predefined subgraph patterns that represent
+common traffic scenarios for isomorphism-based coverage analysis.
+"""
+import networkx as nx
+from graph_creator.models import ActorType
+
+
+def create_lead_vehicle_with_neighbor_vehicle():
+    """Lead vehicle in front with neighbor vehicle."""
+    G = nx.MultiDiGraph()
+    G.add_node("a", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
+    G.add_node("b", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
+    G.add_node("c", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
+    G.add_edge("a", "b", edge_type="following_lead")
+    G.add_edge("b", "a", edge_type="following_lead")
+    G.add_edge("a", "c", edge_type="neighbor_vehicle")
+    G.add_edge("c", "a", edge_type="neighbor_vehicle")
+    return G
+
+
+def create_lead_vehicle_with_neighbor_vehicle_intersection():
+    """Lead vehicle in front with neighbor vehicle at intersection."""
+    G = nx.MultiDiGraph()
+    G.add_node("a", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=True)
+    G.add_node("b", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
+    G.add_node("c", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=True)
+    G.add_edge("a", "b", edge_type="following_lead")
+    G.add_edge("b", "a", edge_type="following_lead")
+    G.add_edge("a", "c", edge_type="neighbor_vehicle")
+    G.add_edge("c", "a", edge_type="neighbor_vehicle")
+    return G
+
+
+def create_lead_vehicle_following_vehicle_in_back():
+    """Lead vehicle in front with following vehicle in the back."""
+    G = nx.MultiDiGraph()
+    G.add_node("a", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
+    G.add_node("b", actor_type=ActorType.VEHICLE, lane_change=False, is_on_intersection=False)
[…]
